---
title: "Testing the Identifying Assumptions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Testing the Identifying Assumptions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

## The identifying assumptions

The RSV estimator relies on two core assumptions:

**Assumption 1 — Stability** (`R ⊥ S | Y`): the conditional distribution of
the RSVs given the outcome is the same in the experimental and observational
samples.  Intuitively, "what R looks like for a unit with Y = y" does not
depend on which study the unit belongs to.

**Assumption 2 — No direct effect** (`R ⊥ D | S_e = 1, Y`): treatment affects
the outcome but does not shift the RSVs *given* the outcome.  If R is a
post-treatment satellite image and D shifts R independently of Y, this
assumption fails.

Both assumptions can be tested directly when Y is observed in the
experimental sample — a setting sometimes called the *validation* design.
In the main use case (Y unobserved in the experimental sample) direct tests
are infeasible; use the J-test as an indirect check instead.

## Setup

```{r}
library(remoteoutcome)
```

We use `sim_rsv_data()` with `n_v = 100` to create 100 overlap units
(`S_e = 1, S_o = 1`) that have both Y and D observed.  The remaining 200
experimental units are purely experimental (`S_e = 1, S_o = 0`) and have
`Y = NA` — the main use case for the estimator.

```{r}
dat <- sim_rsv_data(n_e = 300, n_o = 700, n_v = 100, seed = 42)
R   <- as.matrix(dat[, paste0("R", 1:5)])

# Unit types
table(S_e = dat$S_e, S_o = dat$S_o)
```

## Testing stability

`test_stability()` compares the marginal distribution of each RSV column
conditional on Y between the experimental and observational samples using a
two-sample Kolmogorov-Smirnov test.

For each combination of outcome level `y` and RSV column `R_j`, it tests

> H₀: F(R_j | S = experimental, Y = y) = F(R_j | S = observational, Y = y)

Failure to reject supports stability.

```{r}
stab <- test_stability(R, dat$Y, dat$S_e, dat$S_o)
print(stab)
```

The `p_value` column gives the KS p-value for each (Y level, RSV) pair.
In data generated from our DGP the null holds by construction, so we expect
most p-values to be large.

### What to look for

- **Large p-values** (e.g., > 0.1): no evidence against stability.
- **Small p-values** (e.g., < 0.05): the marginal distribution of that RSV
  differs between samples at that outcome level — a potential violation.
- **Pattern across columns**: a single small p-value may be a false positive;
  many small p-values across columns at the same Y level is more concerning.

### Limitations

The KS test checks the *marginal* distribution of each RSV column, not the
full *joint* conditional.  Low power is expected at small sample sizes.  A
passing test is necessary but not sufficient for stability.

## Testing no direct effect

`test_no_direct_effect()` compares the distribution of each RSV column between
treated and untreated units *within* the experimental sample, conditioning on Y:

> H₀: F(R_j | S = experimental, D = 0, Y = y) = F(R_j | S = experimental, D = 1, Y = y)

```{r}
nde <- test_no_direct_effect(R, dat$Y, dat$D, dat$S_e)
print(nde)
```

Failure to reject (large p-values) supports the no-direct-effect assumption.

### What to look for

- If treatment directly shifts satellite imagery (e.g., a program that changes
  land cover), R distributions will differ between arms even within the same Y
  level — small p-values signal this.
- As with the stability test, isolated small p-values may be chance; a
  systematic pattern across RSV columns is more informative.

## When direct tests are not available

In the typical RSV setting Y is **not observed** in the experimental sample —
that is the whole point of the estimator.  Without an overlap subsample
(`n_v = 0`), both `test_stability()` and `test_no_direct_effect()` will error
with an informative message:

```{r error = TRUE}
dat_no_v <- sim_rsv_data(n_e = 300, n_o = 700, n_v = 0, seed = 42)
R_no_v   <- as.matrix(dat_no_v[, paste0("R", 1:5)])

test_stability(R_no_v, dat_no_v$Y, dat_no_v$S_e, dat_no_v$S_o)
```

```{r error = TRUE}
test_no_direct_effect(R_no_v, dat_no_v$Y, dat_no_v$D, dat_no_v$S_e)
```

### The J-test as an indirect check

When direct tests are infeasible, fit the RSV estimator (using only the purely
experimental and observational units) and use `jtest()`.
It tests whether the efficient (RSV) and naive moment conditions agree on the
same treatment effect — an overidentifying restriction that holds under the
assumptions.

```{r}
fit <- cv.rsv(
  Y   = dat_no_v$Y,
  D   = dat_no_v$D,
  S_e = dat_no_v$S_e,
  S_o = dat_no_v$S_o,
  R   = R_no_v,
  models = list(
    Y   = list(model = "logit"),
    D   = list(model = "logit"),
    S_e = list(model = "logit"),
    S_o = list(model = "logit")
  ),
  nfolds = 5,
  seed   = 42
)

jtest(fit)
```

A large p-value means the two moment conditions are consistent.  Like the KS
tests above, failure to reject is supportive but not conclusive; the J-test has
limited power against specific violations.

## Decision guide

| Situation | Recommended test |
|-----------|-----------------|
| Y observed in experimental sample, single study | `test_stability()` + `test_no_direct_effect()` |
| Y observed in experimental sample, no treatment variation | `test_stability()` only |
| Y **not** observed in experimental sample (main use case) | `jtest()` after fitting `cv.rsv()` |
| Both settings available (validation study) | All three |

## Summary

```{r}
# Quick reference
?test_stability
?test_no_direct_effect
?jtest
```

| Function | Tests | Requires |
|----------|-------|----------|
| `test_stability()` | `R ⊥ S \| Y` | Y observed in both samples |
| `test_no_direct_effect()` | `R ⊥ D \| S_e=1, Y` | Y and D observed in experimental sample |
| `jtest()` | Overidentifying restriction | Fitted `cv.rsv` object, binary Y, no covariates |
