---
title: "Replication: Crop Burning in India"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replication: Crop Burning in India}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

# All computation chunks are skipped if the data cannot be fetched.
# This keeps the vignette CRAN-safe when internet is unavailable.
run_chunks <- tryCatch({
  base_url <- paste0(
    "https://raw.githubusercontent.com/",
    "asheshrambachan/rsv_replication/main/data/clean/cropburn/"
  )
  con <- url(paste0(base_url, "data_obs.csv"))
  close(con)
  TRUE
}, error   = function(e) FALSE,
   warning = function(w) FALSE)
```

## Overview

This vignette replicates the empirical results from Rambachan, Singh, and
Viviano using the **remoteoutcome** package. The application studies the effect
of a crop-burning restriction in India, where:

- **Treatment** `D`: assignment to a fire-restriction zone (0/1).
- **Outcome** `Y`: whether a plot was observed burning in a survey (0/1).
  Observed only in an observational sample of spot-check plots.
- **Remote sensing variable** `R_prob`: a continuous RS-based prediction of
  burning probability, observed for all plots in both samples.
- **Clusters**: villages (`village_id`).

Two samples are available:

| Sample | Description |
|--------|-------------|
| Observational (`data_obs`) | Non-overlapping S_e and S_o: experimental plots have D but not Y; observational plots have Y but not D. |
| Validation (`data_val`) | All plots have D; 444 spot-check plots also have Y (`S_e = 1, S_o = 1`), enabling comparison to a ground-truth ATE. |

The cleaned data are publicly available at
`https://github.com/asheshrambachan/rsv_replication/tree/main/data/clean/cropburn`.

> **Note on bootstrap draws**: The code below uses `B = 100` so the vignette
> runs quickly. The published results use `B = 5,000` (clustered score
> bootstrap, seed 42). With B = 5,000 the estimates are stable to four decimal
> places; with B = 100 expect small Monte Carlo variation of roughly ±0.01 in
> the standard errors.

## Setup

```{r libraries, eval = run_chunks}
library(remoteoutcome)
```

## 1. Load data

```{r load-data, eval = run_chunks}
base_url <- paste0(
  "https://raw.githubusercontent.com/",
  "asheshrambachan/rsv_replication/main/data/clean/cropburn/"
)

data_obs <- read.csv(paste0(base_url, "data_obs.csv"))
data_val <- read.csv(paste0(base_url, "data_val.csv"))
```

`S_e` and `S_o` are derived from the missingness pattern of `D` and `Y`:
a plot is in the experimental sample if `D` is observed, and in the
observational sample if `Y` is observed.

```{r sample-indicators, eval = run_chunks}
# Observational sample: non-overlapping (S = "e" or "o")
S_e_obs <- !is.na(data_obs$D) & !is.na(data_obs$R_prob)
S_o_obs <- !is.na(data_obs$Y) & !is.na(data_obs$R_prob)

# Validation sample: spot-check plots have S = "e,o" (S_e = 1 and S_o = 1)
S_e_val <- !is.na(data_val$D) & !is.na(data_val$R_prob)
S_o_val <- !is.na(data_val$Y) & !is.na(data_val$R_prob)

table(S_e = S_e_obs, S_o = S_o_obs)   # non-overlapping
table(S_e = S_e_val, S_o = S_o_val)   # 444 overlap units
```

## 2. RSV estimator — observational sample

`cv.rsv()` uses 2-fold cross-fitting with logistic regression for all four
nuisance models. Standard errors use the score bootstrap, clustered at the
village level.

```{r rsv-obs, eval = run_chunks}
models <- list(
  Y   = list(model = "logit"),
  D   = list(model = "logit"),
  S_e = list(model = "logit"),
  S_o = list(model = "logit")
)

fit_obs <- cv.rsv(
  Y        = data_obs$Y,
  D        = data_obs$D,
  S_e      = S_e_obs,
  S_o      = S_o_obs,
  R        = data_obs$R_prob,
  y_levels = c(0L, 1L),
  models   = models,
  nfolds   = 2,
  clusters = data_obs$village_id,
  seed     = 42L
)

fit_obs <- add_se(
  fit_obs,
  method      = "score_bootstrap",
  B           = 100,
  clusters    = data_obs$village_id,
  seed        = 42L
)

summary(fit_obs)
```

Published results (B = 5,000): RSV = **0.1417** (SE: 0.0782),
Naive = **0.1929** (SE: 0.0989).

## 3. RSV estimator — validation sample

The validation sample has 444 spot-check plots with `S_e = 1` and `S_o = 1`.
These overlap units enable direct assumption tests
(see Section 7 and `vignette("assumption-testing")`).

```{r rsv-val, eval = run_chunks}
fit_val <- cv.rsv(
  Y        = data_val$Y,
  D        = data_val$D,
  S_e      = S_e_val,
  S_o      = S_o_val,
  R        = data_val$R_prob,
  y_levels = c(0L, 1L),
  models   = models,
  nfolds   = 2,
  clusters = data_val$village_id,
  seed     = 42L
)

fit_val <- add_se(
  fit_val,
  method      = "score_bootstrap",
  B           = 100,
  clusters    = data_val$village_id,
  seed        = 42L
)

summary(fit_val)
```

Published results (B = 5,000): RSV = **0.1210** (SE: 0.0684),
Naive = **0.1565** (SE: 0.0803).

## 4. J-test for overidentifying restrictions

The J-test checks whether the efficient (RSV) and naive moment conditions
agree — an indirect test of the stability and no-direct-effect assumptions.

```{r jtest, eval = run_chunks}
jtest(fit_obs)
jtest(fit_val)
```

Published results: J = 1.744 (p = 0.187) for the observational sample;
J = 1.324 (p = 0.250) for the validation sample. Neither rejects the null;
the two moment conditions are mutually consistent.

## 5. Relevance

Relevance measures how informative `R_prob` is for identifying
treatment-induced variation in `Y`.

```{r relevance, eval = run_chunks}
relevance(fit_obs, level = 0.90)
relevance(fit_val, level = 0.90)
```

Published results (90% CI):
observational — 0.2622 (SE: 0.0480, CI: [0.1833, 0.3411]);
validation — 0.3157 (SE: 0.0533, CI: [0.2280, 0.4035]).

Relevance is substantially positive in both samples, confirming that
`R_prob` carries useful information about `Y`.

## 6. Summary table

The table below replicates Table 2 from the paper, comparing the RSV and
naive estimates across samples.

```{r summary-table, echo = FALSE, eval = TRUE}
res <- data.frame(
  Estimator = c(
    "RSV (optimal representation)", "",
    "Naive (simple representation)", "",
    "J statistic",
    "p-value"
  ),
  Validation = c(
    "0.1210", "(0.0684)",
    "0.1565", "(0.0803)",
    "1.3242", "0.2498"
  ),
  Observational = c(
    "0.1417", "(0.0782)",
    "0.1929", "(0.0989)",
    "1.7439", "0.1866"
  )
)
knitr::kable(
  res,
  align    = c("l", "r", "r"),
  col.names = c("", "Validation", "Observational"),
  caption  = "Treatment effect estimates (standard errors in parentheses). Published results, B = 5,000."
)
```

## 7. Assumption tests (validation sample only)

The validation sample has overlap units (`S_e = 1, S_o = 1`), so we can
directly test the identifying assumptions. These tests are not possible for
the observational sample because `Y` is never observed for experimental units
there.

```{r assumption-tests, eval = run_chunks}
R_mat <- matrix(data_val$R_prob)

# Stability: F(R | S=e, Y=y) = F(R | S=o, Y=y)
test_stability(R_mat, data_val$Y, S_e_val, S_o_val)

# No direct effect: F(R | S=e, D=0, Y=y) = F(R | S=e, D=1, Y=y)
test_no_direct_effect(R_mat, data_val$Y, data_val$D, S_e_val)
```

Failure to reject in both tests supports the identifying assumptions.

## Reproducing the published results

To exactly match the published standard errors, increase `B` to 5,000 and
use 7 parallel threads:

```{r full-replication, eval = FALSE}
fit_obs <- add_se(
  fit_obs,
  method      = "score_bootstrap",
  B           = 5000,
  clusters    = data_obs$village_id,
  seed        = 42L,
  num.threads = 7
)
```

The full replication (both samples, B = 5,000) takes approximately 10–15
minutes on a modern laptop with 7 threads.
