---
title: "Getting Started with remoteoutcome"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with remoteoutcome}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

## Overview

**remoteoutcome** estimates average treatment effects (ATEs) when the outcome
is observed only in an observational sample but not in the experimental sample.
The key idea is to use *remotely sensed variables* (RSVs) — covariates observed
in both samples that are informative about the outcome — to bridge the two
samples.

This vignette walks through the full workflow on a simulated binary-outcome
dataset:

1. Simulate data with `sim_rsv_data()`
2. Estimate the ATE with `cv.rsv()` (cross-fitting)
3. Add standard errors with `add_se()`
4. Inspect the fit: coefficients, relevance, and the J-test

## Setup

```{r}
library(remoteoutcome)
```

## 1. Data structure

The package expects a stacked dataset containing both the experimental and
observational samples.  Each row is a unit; the two samples are distinguished
by indicator columns `S_e` (experimental) and `S_o` (observational).

| Column | Pure experimental (`S_e=1, S_o=0`) | Overlap (`S_e=1, S_o=1`) | Pure observational (`S_e=0, S_o=1`) |
|--------|:----------------------------------:|:------------------------:|:-----------------------------------:|
| `Y`    | **NA**                             | observed                 | observed                            |
| `D`    | observed (0/1)                     | observed (0/1)           | **NA**                              |
| `R`    | observed                           | observed                 | observed                            |

`sim_rsv_data()` generates data from this design.  The true ATE is controlled
by `tau` (default 0.10).  Overlap units (`S_e = 1 & S_o = 1`) are created by
setting `n_v > 0`; they have both Y and D observed and enable direct assumption
tests (see `vignette("assumption-testing")`).

```{r}
dat <- sim_rsv_data(n_e = 300, n_o = 700, tau = 0.10, seed = 42)
head(dat)
```

```{r}
# Outcome is NA for purely experimental units (S_e=1, S_o=0)
table(S_e = dat$S_e, S_o = dat$S_o, Y_observed = !is.na(dat$Y))
```

## 2. Fitting the RSV estimator

`cv.rsv()` uses K-fold cross-fitting: models are fit on held-out folds to avoid
overfitting the predictions used in the moment conditions.

The `models` argument specifies the learner for each nuisance component.
`"logit"` (logistic regression) is fast and works well for small-to-medium
samples.  For larger datasets or complex relationships, use
`list(model = "rf", num.trees = 500)` to fit a random forest.

```{r}
R <- as.matrix(dat[, paste0("R", 1:5)])

fit <- cv.rsv(
  Y   = dat$Y,
  D   = dat$D,
  S_e = dat$S_e,
  S_o = dat$S_o,
  R   = R,
  models = list(
    Y   = list(model = "logit"),
    D   = list(model = "logit"),
    S_e = list(model = "logit"),
    S_o = list(model = "logit")
  ),
  nfolds = 5,
  seed   = 42
)

print(fit)
```

The `coefficients` field contains the estimated ATE.  The true value is 0.10;
the RSV estimate should be close.

```{r}
fit$coefficients
```

## 3. Adding standard errors

Standard errors are computed via the influence-function approach by default,
which is fast and suitable for most use cases.

```{r}
fit <- add_se(fit, method = "influence", B = 500, seed = 42)
summary(fit)
```

`score_bootstrap` re-draws influence scores rather than refitting the full
model, giving a quick bootstrap approximation:

```{r eval = FALSE}
fit <- add_se(fit, method = "score_bootstrap", B = 1000, seed = 42)
```

For fully non-parametric uncertainty estimates, `method = "bootstrap"` refits
`cv.rsv()` on each bootstrap sample (slow but most general):

```{r eval = FALSE}
fit <- add_se(fit, method = "bootstrap", B = 200, seed = 42)
```

## 4. Relevance

*Relevance* measures how strongly the RSVs predict treatment-induced variation
in the outcome.  Higher relevance indicates more informative RSVs and tighter
estimates.

```{r}
relevance(fit)
```

The table shows the RSV estimator alongside the naive estimator (which ignores
the experimental sample structure).  The RSV relevance is typically higher
when the nuisance models are well-specified.

## 5. J-test

The J-test checks whether the efficient (RSV) and naive moment conditions
agree on the same treatment effect — an indirect test of the identifying
assumptions (stability and no direct effect).

```{r}
jtest(fit)
```

A large p-value means the two moment conditions are consistent, which
supports the assumptions.  A small p-value suggests misspecification in the
nuisance models or a violation of an assumption.

## 6. Clustered data

When observations are grouped (e.g., villages, schools), pass cluster IDs to
`cv.rsv()` so the train/test split respects cluster boundaries, and to
`add_se()` for cluster-robust standard errors:

```{r eval = FALSE}
cluster_id <- rep(1:100, each = 10)  # 100 clusters of 10

fit_cl <- cv.rsv(
  Y = dat$Y, D = dat$D, S_e = dat$S_e, S_o = dat$S_o, R = R,
  models  = list(
    Y   = list(model = "logit"),
    D   = list(model = "logit"),
    S_e = list(model = "logit"),
    S_o = list(model = "logit")
  ),
  clusters = cluster_id,
  nfolds   = 5,
  seed     = 42
)

fit_cl <- add_se(fit_cl, method = "bootstrap", B = 200,
                 clusters = cluster_id, seed = 42)
```

## Summary

| Function        | Purpose                                              |
|-----------------|------------------------------------------------------|
| `sim_rsv_data()`| Simulate experimental + observational data           |
| `cv.rsv()`      | Estimate ATE via cross-fitted RSV estimator          |
| `rsv_split()`   | Estimate ATE via single train/test split             |
| `add_se()`      | Attach standard errors to a fitted object            |
| `relevance()`   | Extract RSV and naive relevance estimates            |
| `jtest()`       | J-test for overidentifying restrictions              |

See `vignette("assumption-testing")` for how to test the stability and
no-direct-effect assumptions when the data allow it.
