---
title: "Replication: Smartcards in India"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replication: Smartcards in India}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

# All computation chunks are skipped if the data cannot be fetched.
# The file is stored with Git LFS; download.file() is required because
# read.csv(url(...)) does not follow the LFS redirect.
.lfs_url <- paste0(
  "https://media.githubusercontent.com/media/",
  "asheshrambachan/rsv_replication/main/data/clean/smartcards/data.csv"
)
.tmp_sc <- tempfile(fileext = ".csv")
run_chunks <- tryCatch({
  status <- download.file(.lfs_url, .tmp_sc, quiet = TRUE, method = "libcurl")
  status == 0L && file.exists(.tmp_sc) && file.size(.tmp_sc) > 1000L
}, error = function(e) FALSE)
```

## Overview

This vignette replicates the smartcards empirical analysis and simulation study
from Rambachan, Singh, and Viviano using the **remoteoutcome** package.
The application re-analyses the effect of biometric payment infrastructure
(Muralidharan et al. 2016) on three village-level binary outcomes in Andhra
Pradesh, India:

| Outcome | Description |
|---------|-------------|
| `Ycons` | Consumption above median |
| `Ylowinc` | Low-income household |
| `Ymidinc` | Middle-income household |

The study design involves four waves mapped to sample indicators:

| Wave | `S_e` | `S_o` | D observed | Y observed |
|------|:-----:|:-----:|:----------:|:----------:|
| Experimental: Treated (2010) | 1 | 0 | ✓ | — |
| Experimental: Untreated (2011) | 1 | 1 | ✓ | ✓ |
| Experimental: Untreated (2012) | 1 | 0 | ✓ | — |
| Observational (N/A) | 0 | 1 | — | ✓ |

The 2011 untreated units are **overlap units** (`S_e = S_o = 1`): both treatment
and outcome are observed, enabling direct assumption tests.

The remote sensing variable is `R_PC1_scaled`, the first principal component
of MOSAIKS satellite features (pre-computed). The published results use 1,000
satellite PCs and nighttime luminosity with random forests; here we use a
single PC with logistic nuisance models for speed. A full-replication template
(RF, all features, `B = 500`) is given in Section 8.

> **Note on computation**: Section 3 uses `rsv()` with RF nuisance models
> (luminosity + 100 satellite PCs, 100 trees, depth 10) and score-bootstrap SEs
> (`B = 50`, clustered by geography). Sections 5–6 use `cv.rsv()` with
> logistic nuisance models, influence-function SEs, and `B = 50` Monte Carlo
> replications. The published analysis uses `B = 1,000`, `B_se = 100`,
> 1,000 trees, and 4,000 satellite PCs + luminosity; vignette estimates are
> therefore simplified approximations. Results are shown for `Ymidinc` only.

## Setup

```{r libraries, eval = run_chunks}
library(remoteoutcome)
```

Helper functions shared with the Uganda forest cover vignette.

```{r helpers, eval = run_chunks}
# KL-tilting: find p1 s.t. E_{p1}[y] = E_{p0}[y] + tau
compute_p1_kl <- function(p0, y, tau, tol = 1e-10) {
  mu_target <- sum(y * p0) + tau
  support <- p0 > 0
  p0s <- p0[support]; ys <- y[support]
  if (abs(tau) < tol) return(p0)
  f <- function(lam) {
    z <- lam * ys - max(lam * ys)
    w <- p0s * exp(z); w <- w / sum(w)
    sum(ys * w) - mu_target
  }
  lam <- uniroot(f, c(-100, 100), tol = tol)$root
  z <- lam * ys - max(lam * ys); w <- p0s * exp(z); w <- w / sum(w)
  p1 <- numeric(length(p0)); p1[support] <- w; p1
}

sample_Y_exp <- function(D, y_levels, p0, p1) {
  K <- length(y_levels); out <- integer(length(D))
  idx0 <- D == 0; idx1 <- D == 1
  if (any(idx0)) out[idx0] <- sample.int(K, sum(idx0), replace = TRUE, prob = p0)
  if (any(idx1)) out[idx1] <- sample.int(K, sum(idx1), replace = TRUE, prob = p1)
  y_levels[out]
}

sample_R_given_Y <- function(R_given_Y, y_vals, y_levels) {
  out <- matrix(NA_real_, length(y_vals), ncol(R_given_Y[[1]]))
  for (k in seq_along(y_levels)) {
    idx <- which(y_vals == y_levels[k])
    if (!length(idx)) next
    draw <- sample.int(nrow(R_given_Y[[k]]), length(idx), replace = TRUE)
    out[idx, ] <- R_given_Y[[k]][draw, , drop = FALSE]
  }
  out
}

sim_noexp <- function(n_e, n_o, y_levels, p0, p1, alpha_o, R_given_Y) {
  D_e <- rbinom(n_e, 1, 0.5)
  Y_e <- sample_Y_exp(D_e, y_levels, p0, p1)
  R_e <- sample_R_given_Y(R_given_Y, Y_e, y_levels)
  Y_e <- rep(NA_real_, n_e)
  K   <- length(y_levels)
  Y_o <- y_levels[sample.int(K, n_o, replace = TRUE, prob = alpha_o)]
  R_o <- sample_R_given_Y(R_given_Y, Y_o, y_levels)
  list(Y = c(Y_e, Y_o), D = c(D_e, rep(NA_real_, n_o)),
       R = rbind(R_e, R_o),
       S_e = c(rep(1L, n_e), rep(0L, n_o)),
       S_o = c(rep(0L, n_e), rep(1L, n_o)))
}

summarise_sims <- function(draws, tau) {
  ci_mult <- qnorm(0.975)
  bias <- mean(draws$tauhat - tau, na.rm = TRUE)
  sd_e <- sd(draws$tauhat, na.rm = TRUE)
  cov  <- mean(tau >= draws$tauhat - ci_mult * draws$se &
               tau <= draws$tauhat + ci_mult * draws$se, na.rm = TRUE)
  data.frame(tau             = tau,
             normalized_bias = round(bias / sd_e, 3),
             coverage        = round(cov, 3),
             rmse            = round(sqrt(mean((draws$tauhat - tau)^2, na.rm = TRUE)), 4),
             avg_se          = round(mean(draws$se, na.rm = TRUE), 4),
             n_sims          = sum(!is.na(draws$tauhat)))
}
```

## 1. Load data

The cleaned dataset is stored with Git LFS and was downloaded to a temp file
during the setup check.

```{r load-data, eval = run_chunks}
dt <- read.csv(.tmp_sc)

cat("Rows:", nrow(dt), "\n")
print(table(wave = dt$wave))
```

Derive sample indicators and mask Y/D for the RSV estimator.

```{r indicators, eval = run_chunks}
OUTCOMES <- "Ymidinc"

# S_e = all experimental waves; S_o = 2011 untreated (overlap) + observational
dt$S_e <- as.integer(dt$wave != "Observational (N/A)")
dt$S_o <- as.integer(dt$wave %in% c("Experimental: Untreated (2011)",
                                     "Observational (N/A)"))

# Mask: D unobserved for observational; Y unobserved for pure experimental
dt$D_est <- ifelse(dt$S_e == 1L, dt$D, NA_real_)
for (yv in OUTCOMES) {
  dt[[paste0(yv, "_est")]] <- ifelse(dt$S_o == 1L, dt[[yv]], NA_real_)
}

cat("S_e=1:", sum(dt$S_e), " S_o=1:", sum(dt$S_o),
    " Overlap (S_e=S_o=1):", sum(dt$S_e == 1L & dt$S_o == 1L), "\n")
```

## 2. Assumption tests

Both tests use `R_PC1_scaled` (the pre-computed first satellite PC) and the
raw (unmasked) Y and D values, consistent with the published analysis.

### Stability

`test_stability()` tests H₀: F(R | S = experimental, D = 0, Y = y) =
F(R | S = observational, D = 0, Y = y). Conditioning on D = 0 avoids
conflating the sensing mechanism with the treatment effect.

```{r stability, eval = run_chunks}
R_pc1 <- matrix(dt$R_PC1_scaled, ncol = 1L, dimnames = list(NULL, "R_PC1"))

# Subset to D = 0; observational units have D recorded in the raw data
d0 <- !is.na(dt$D) & dt$D == 0L
S_e_d0 <- as.integer(dt$S_e[d0] == 1)
S_o_d0 <- as.integer(dt$S_o[d0] == 1)

for (yv in OUTCOMES) {
  cat(sprintf("\n--- Stability: %s ---\n", yv))
  print(test_stability(R_pc1[d0, , drop = FALSE], dt[[yv]][d0], S_e_d0, S_o_d0))
}
```

### No direct effect

`test_no_direct_effect()` tests H₀: F(R | S = experimental, D = 0, Y = y) =
F(R | S = experimental, D = 1, Y = y). Uses all experimental units; Y is
taken from the raw data (all waves have Y recorded before masking).

```{r no-direct-effect, eval = run_chunks}
exp_mask <- dt$S_e == 1L
R_exp    <- R_pc1[exp_mask, , drop = FALSE]
S_e_exp  <- rep(1L, sum(exp_mask))

for (yv in OUTCOMES) {
  cat(sprintf("\n--- No direct effect: %s ---\n", yv))
  print(test_no_direct_effect(R_exp, dt[[yv]][exp_mask],
                               dt$D[exp_mask], S_e_exp))
}
```

No systematic pattern of small p-values across outcomes or Y levels; failure
to reject supports the no-direct-effect assumption.

## 3. Empirical RSV estimates

`rsv()` fits nuisance models on the full sample and returns efficient and naive
estimates simultaneously. Standard errors are clustered score bootstrap
(`B = 50`, clustered by geographic unit). Random forest nuisance models use
nighttime luminosity and the first 100 satellite PCs; the published analysis
uses 4,000 satellite PCs and luminosity with larger forests.

```{r empirical-R, eval = run_chunks}
lum_cols <- grep("^luminosity_", names(dt), value = TRUE)
sat_cols <- paste0("satellite_", 1:100)
R_est    <- as.matrix(dt[, c(lum_cols, sat_cols)])
```

```{r empirical-rsv, eval = run_chunks}
rf_spec <- list(model = "rf", num.trees = 100, min.node.size = 20,
                max.depth = 10, verbose = FALSE)
models  <- list(Y = rf_spec, D = rf_spec, S_e = rf_spec, S_o = rf_spec)

n_threads <- max(1L, parallel::detectCores() - 1L)

fits <- list()
for (yv in OUTCOMES) {
  fit <- rsv(
    Y        = dt[[paste0(yv, "_est")]],
    D        = dt$D_est,
    S_e      = dt$S_e,
    S_o      = dt$S_o,
    R        = R_est,
    y_levels = c(0L, 1L),
    yK       = 0L,
    models   = models,
    seed     = 42L,
    num.threads = n_threads
  )
  fits[[yv]] <- add_se(fit, clusters = dt$clusters,
                       method = "score_bootstrap", B = 50L,
                       seed = 42L, num.threads = n_threads)
}

for (yv in OUTCOMES) {
  f <- fits[[yv]]
  cat(sprintf("%s:  RSV = %7.4f (SE: %.4f)\n", yv, f$coefficients, f$se))
}
```

### Relevance

```{r relevance, eval = run_chunks}
for (yv in OUTCOMES) {
  cat(sprintf("\n--- Relevance: %s ---\n", yv)); print(relevance(fits[[yv]]))
}
```

### Comparison to published results

The published estimates use 4,000 satellite PCs and nighttime luminosity with
random forests (`B = 1,000` score bootstrap), which capture richer variation
in welfare outcomes than a single PC. The table below reports the published
point estimates for reference.

```{r empirical-published, echo = FALSE, eval = TRUE}
pub_emp <- data.frame(
  Outcome      = "Ymidinc",
  RSV          = "-0.0450",
  RSV_SE       = "(0.0360)",
  Relevance    = "0.453",
  Relevance_SE = "(0.049)"
)
knitr::kable(pub_emp,
  col.names = c("Outcome", "RSV", "(SE)", "Relevance", "(SE)"),
  caption = paste("Published empirical results (full sample, RF + 4,000 satellite",
                  "PCs + luminosity, B = 1,000 score bootstrap)."))
```

Relevance is well above zero, confirming the satellite features carry useful
information about welfare outcomes.

## 4. Simulation setup

The simulation DGP is calibrated to the smartcards data. The baseline outcome
distribution `p0` = `alpha_e` uses only the **untreated** experimental waves
(2011 and 2012), so that the treated-wave outcomes do not inflate the control
distribution. The treated potential outcome distribution `p1` is obtained by
KL-tilting `p0` to achieve a mean shift of `tau`.

```{r dgp-setup, eval = run_chunks}
y_levels <- c(0L, 1L)

# Untreated experimental waves only for alpha_e calibration
S_e_sim <- dt$wave %in% c("Experimental: Untreated (2011)",
                           "Experimental: Untreated (2012)")
S_o_sim <- dt$wave == "Observational (N/A)"

# Per-outcome DGP parameters
dgp <- lapply(setNames(OUTCOMES, OUTCOMES), function(yv) {
  Y_all   <- dt[[yv]]
  p0      <- as.numeric(prop.table(table(Y_all[S_e_sim])))
  alpha_o <- as.numeric(prop.table(table(Y_all[S_o_sim])))
  # R|Y lookup from all units with Y observed (R_PC1_scaled, single column)
  R_given_Y <- lapply(y_levels, function(y)
    matrix(dt$R_PC1_scaled[!is.na(Y_all) & Y_all == y], ncol = 1L))
  list(p0 = p0, alpha_o = alpha_o, R_given_Y = R_given_Y)
})

TAU_GRID <- c(0, 0.05, 0.10, 0.15, 0.20)
B        <- 50L
SEED     <- 42L
N_E      <- 3000L
N_O      <- 1000L   # published also runs n_o = 500 and 1500

cat("DGP calibration (p0 = alpha_e for untreated experimental units):\n")
for (yv in OUTCOMES) {
  cat(sprintf("  %s: p0 = [%.3f, %.3f]  alpha_o = [%.3f, %.3f]\n",
              yv, dgp[[yv]]$p0[1], dgp[[yv]]$p0[2],
              dgp[[yv]]$alpha_o[1], dgp[[yv]]$alpha_o[2]))
}
```

## 5. No experimental outcomes simulation

Y is entirely unobserved in the experimental sample. The simulation is run for
all three outcomes with `B = 50` Monte Carlo replications and influence-function
SEs.

```{r sim-run, eval = run_chunks}
run_noexp_sc <- function(yv, tau, b) {
  set.seed(SEED + b)
  p1  <- compute_p1_kl(dgp[[yv]]$p0, y_levels, tau)
  dat <- sim_noexp(N_E, N_O, y_levels, dgp[[yv]]$p0, p1,
                   dgp[[yv]]$alpha_o, dgp[[yv]]$R_given_Y)
  fit <- cv.rsv(Y = dat$Y, D = dat$D, S_e = dat$S_e, S_o = dat$S_o, R = dat$R,
                y_levels = y_levels, yK = y_levels[1], models = models,
                nfolds = 2L, seed = NULL, num.threads = 1L)
  fit <- add_se(fit, method = "influence")
  data.frame(outcome = yv, tau = tau, tauhat = fit$coefficients, se = fit$se)
}

cat("Running simulation (3 outcomes x 5 tau values x B=50)...\n")
sim_results <- do.call(rbind, lapply(OUTCOMES, function(yv) {
  do.call(rbind, lapply(TAU_GRID, function(tau) {
    cat(sprintf("  %s  tau=%.2f\n", yv, tau))
    do.call(rbind, lapply(seq_len(B), run_noexp_sc, yv = yv, tau = tau))
  }))
}))
```

```{r sim-summary, eval = run_chunks}
sim_tbl <- do.call(rbind, lapply(OUTCOMES, function(yv) {
  cbind(outcome = yv,
        do.call(rbind, lapply(TAU_GRID, function(tv)
          summarise_sims(sim_results[sim_results$outcome == yv &
                                     sim_results$tau == tv, ], tv))))
}))
print(sim_tbl, row.names = FALSE)
```

## 6. Comparison to published simulation results

Published results use `B = 500`, `score_bootstrap` with `B_se = 100`,
`n_e = 3,000`, `n_o = 1,000`, RF nuisance models with 1,000 trees, and 1,000
satellite PCs plus luminosity as RSVs. With `B = 50` and a single-PC logit
specification, coverage SEs are ≈ ±0.03 and average SEs may differ due to the
simplified RSV.

```{r sim-published, echo = FALSE, eval = TRUE}
pub_sim <- data.frame(
  tau       = c(0.00, 0.05, 0.10, 0.15, 0.20),
  norm_bias = c(0.067,  0.003,  0.074,  0.144,  0.106),
  coverage  = c(0.970,  0.954,  0.962,  0.954,  0.938),
  rmse      = c(0.0505, 0.0548, 0.0553, 0.0622, 0.0670),
  avg_se    = c(0.0522, 0.0543, 0.0570, 0.0590, 0.0619)
)
knitr::kable(
  pub_sim,
  digits    = c(2, 3, 3, 4, 4),
  col.names = c("τ", "Norm. bias", "Coverage", "RMSE", "Avg SE"),
  caption   = paste("Published RSV simulation results (Ymidinc): n_e = 3,000, n_o = 1,000,",
                    "B = 500, B_se = 100, RF + 1,000 satellite PCs + luminosity.")
)
```

Near-zero normalized bias and ~96–97% coverage across τ values confirm that
`cv.rsv()` is well-calibrated for `Ymidinc` under this DGP.

## 7. Reproducing the published results

To exactly match the published estimates, use random forest nuisance models
with the full satellite feature set and `score_bootstrap` SEs. The empirical
analysis uses all 4,000 satellite PCs; the simulation uses the first 1,000.

```{r full-replication, eval = FALSE}
# Empirical: luminosity + all 4,000 satellite PCs
R_emp <- as.matrix(dt[, c(grep("^luminosity_", names(dt), value = TRUE),
                           paste0("satellite_", 1:4000))])

# Simulation: luminosity + first 1,000 satellite PCs
R_sim <- as.matrix(dt[, c(grep("^luminosity_", names(dt), value = TRUE),
                           paste0("satellite_", 1:1000))])

rf_spec   <- list(model = "rf", num.trees = 1000L,
                  min.node.size = 20L, max.depth = 10L, verbose = FALSE)
models_rf <- list(Y = rf_spec, D = rf_spec, S_e = rf_spec, S_o = rf_spec)

fits_full <- list()
for (yv in OUTCOMES) {
  fit <- rsv(
    Y        = dt[[paste0(yv, "_est")]],
    D        = dt$D_est,
    S_e      = dt$S_e,
    S_o      = dt$S_o,
    R        = R_emp,
    y_levels = c(0L, 1L),
    yK       = 0L,
    models   = models_rf,
    seed     = 42L,
    num.threads = parallel::detectCores() - 1L
  )
  fits_full[[yv]] <- add_se(fit,
    method      = "score_bootstrap",
    B           = 1000L,
    clusters    = dt$clusters,
    seed        = 42L,
    num.threads = parallel::detectCores() - 1L
  )
}

# Full simulation (B = 500, RF, 1,000 satellite PCs)
run_full_sc <- function(yv, tau, b) {
  set.seed(SEED + b)
  R_given_Y_sim <- lapply(y_levels, function(y)
    R_sim[!is.na(dt[[yv]]) & dt[[yv]] == y, , drop = FALSE])
  p1  <- compute_p1_kl(dgp[[yv]]$p0, y_levels, tau)
  dat <- sim_noexp(N_E, N_O, y_levels, dgp[[yv]]$p0, p1,
                   dgp[[yv]]$alpha_o, R_given_Y_sim)
  fit <- cv.rsv(Y = dat$Y, D = dat$D, S_e = dat$S_e, S_o = dat$S_o, R = dat$R,
                y_levels = y_levels, yK = y_levels[1], models = models_rf,
                nfolds = 2L, seed = NULL, num.threads = 1L)
  fit <- add_se(fit, method = "score_bootstrap", B = 100L,
                seed = NULL, num.threads = 1L)
  data.frame(outcome = yv, tau = tau, tauhat = fit$coefficients, se = fit$se)
}
```

The full empirical run takes approximately 30–60 minutes; the full simulation
(`B = 500`) takes several hours with parallel cores.
