---
title: "Estimating Treatment Effects with Remotely Sensed Variables: Original Predictions"
author: "Rambachan, Singh, and Viviano (2025)"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Estimating Treatment Effects with Remotely Sensed Variables: Original Predictions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Overview

This vignette demonstrates how to use the `remoteoutcome` package to estimate treatment effects when outcomes are measured using remotely sensed variables (RSVs). Specifically, it replicates **Figure 7**  of *Rambachan, Singh, and Viviano (2025)* using the original `randomForest`-based predictions from the paper and applies the RSV estimator using those precomputed values. It complements `vignette("poverty-application")`, which retrains all predictive models using the `ranger` package for improved computational efficiency and reproducibility. 

Because the same estimator and resampling procedure are used, the treatment effect estimates here are **identical** to those in *Rambachan, Singh, and Viviano (2025)*. There is a minimal variation in confidence intervals that arise from the updates in bootstrap procedure, not from the underlying estimator or data. Specifically, the paper computed the bootstrap standard errors using the `boot` package with parallelization without setting the seed. We now use a *custom cluster bootstrap* that supports deterministic seeding via `fix_seed = TRUE`, setting `set.seed(b)` for the *b*‑th replication. This preserves the resampling logic, ensuring reproducibility even under parallel execution.


# Data

The dataset combines experimental and remotely sensed information used in the poverty application.  
It includes:

- **Outcome variables:** Poverty indicators such as household consumption and income measures.  
- **Treatment variable:** Program assignment indicating whether a village received the Smartcard intervention.  
- **Remotely sensed variables:** Satellite-based measures of luminosity and related features.  
- **Sample indicators:** Experimental design identifiers distinguishing treatment and control groups.

```{r load-data}
library(psych)
library(latex2exp)
library(stringr)
library(knitr)
library(tibble)
library(kableExtra)
library(dplyr, warn.conflicts = F)
library(ggplot2, warn.conflicts = F)
library(fixest)
library(tidyr)
library(remoteoutcome)

cores <- 7

# Load poverty data (included in package)
# data
# data_real
# data_synth
```

```{r tab-data, echo=FALSE}
data %>% 
  select(shrid2, clusters, wave, D, Ylowinc, Ymidinc, Ycons, luminosity_max.2012, luminosity_max.2013, satellite_1, satellite_2) %>%
  group_by(D) %>%
  slice(1:2) %>%
  ungroup() %>%
  head()  %>%
  kable(digits = 3, caption="Data overview") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r}
cat(sprintf("Total observations: %d\n", nrow(data)))
cat(sprintf("Number of clusters (subdistricts): %d\n",
            length(unique(data$clusters))))
```

The poverty outcomes analyzed in the application are defined as follows:

  - `Ycons`: Household consumption below 25th percentile.
  - `Ylowinc`: No household members earning more than 5,000 INR per month.
  - `Ymidinc`: No household members earning more than 10,000 INR per month.
  
```{r tab-outcomes, echo=FALSE}
data %>% 
  select(Ycons, Ylowinc, Ymidinc) %>%
  rename(c("Consumption"=Ycons, "Low Income"=Ylowinc, "Middle Income"=Ymidinc)) %>%
  describe() %>%
  select(c("n", "mean", "sd", "min", "max")) %>%
  rownames_to_column(var = "Outcome") %>%
  kable(digits = 3, caption="Outcome Variables") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

The experiment was implemented in multiple phases, with villages randomly assigned to receive Smartcards in different years:

- **Experimental: Treated (2010)**: Villages that received Smartcards in 2010
- **Experimental: Untreated (2011)**: Buffer villages assigned to receive Smartcards in 2011.
- **Experimental: Untreated (2012)**: Control villages assigned to receive Smartcards in 2012.
- **Observational (N/A)**: Non-study or holdout villages excluded from the experiment.

```{r tab-exp-design, echo=FALSE}
tab <- data %>% 
  group_by(`Sample (Smartcard)`) %>%
  summarise(
    `Number of villages`        = n_distinct(shrid2),
    `Average population`        = round(mean(tot_p), 0),
    `Average fraction female`   = round(mean(tot_f), 3),
    .groups = "drop"
  ) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column(var = "Metric")

## Use first row as column names
colnames(tab) <- tab[1, ]
tab <- tab[-1, ]

## Create markdown table
kable(tab, row.names = FALSE, caption="Experimental Design") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Analysis

## RSV Estimation with Real Sample Definition

We demonstrates how to estimate treatment effects using the **RSV estimator** under the *real sample definition*.

Following the study design, we construct two key sample indicators:

- **Experimental sample** (\(S_e\)): Villages that were part of the randomized evaluation
  - Experimental: Treated (2010)
  - Experimental: Untreated (2011)
  - Experimental: Untreated (2012)
  
- **Observational sample** (\(S_o\)): 
  - Experimental: Untreated (2011)
  - Observational (N/A)

This creates an overlap where *Experimental: Untreated (2011)* villages are in both samples.

```{r tab-realS, echo = FALSE}
# Verify sample structure
data_real %>%
  group_by(S) %>%
  summarise(
    `D = 0` = sum(D == 0, na.rm = TRUE),
    `D = 1` = sum(D == 1, na.rm = TRUE),
    `D = NA` = sum(is.na(D)),
    `Ycons = 0` = sum(Ycons == 0, na.rm = TRUE),
    `Ycons = 1` = sum(Ycons == 1, na.rm = TRUE),
    `Ycons = NA` = sum(is.na(Ycons)),
    `Ylowinc = 0` = sum(Ylowinc == 0, na.rm = TRUE),
    `Ylowinc = 1` = sum(Ylowinc == 1, na.rm = TRUE),
    `Ylowinc = NA` = sum(is.na(Ylowinc)),
    `Ymidinc = 0` = sum(Ymidinc == 0, na.rm = TRUE),
    `Ymidinc = 1` = sum(Ymidinc == 1, na.rm = TRUE),
    `Ymidinc = NA` = sum(is.na(Ymidinc))
  ) %>%
  pivot_longer(cols = -S, names_to = "Variable", values_to = "Count") %>%
  pivot_wider(names_from = S, values_from = Count) %>%
  kable(format = "markdown", digits = 3, caption = "Distribution of D and Y by S (Real Definition)", col.names = c("Varibale & Value", "Experimental & Observational Sample", "Experimental Sample Only", "Observational Sample Only")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

We now estimate treatment effects for each poverty outcome using the `rsv_estimate()` function.
This function implements the RSV estimator described in **Rambachan, Singh, and Viviano (2025)**.

In this vignette, instead of retraining predictive models, we use pre-computed predictions generated with the `randomForest` package--the same method used to produce the estimates reported in the paper.
This approach allows the results here to closely mirror the published values. There is still minor differences in CI due to changes in bootstrap procedure as mentioned previously.

Key arguments in this example:

- `Y`, `D`, `S_e`, `S_o`: True observed values of outcomes, treatment assignment, and sample indicators.
- `pred_Y`, `pred_D`, `pred_S_e`, `pred_S_o`: Predictions generated using the original `randomForest` models.
- `method = "none"`: Uses all available data for both training and estimation (no cross-fitting).
- `se = TRUE`: Requests bootstrap-based standard errors with cluster resampling.
- `se_params`: Controls bootstrap settings:
  - `B = 1000`: Number of bootstrap replications.
  - `clusters`: Subdistrict-level cluster identifiers.
  - `fix_seed = TRUE`: Ensures deterministic seeding for reproducibility.
  - `cores`: Number of cores used for parallel bootstrap computation.

```{r rsv-real-Ycons, message=FALSE}
# Consumption prediction data (included in package)
# pred_real_Ycons
rsv_real_Ycons <- rsv_estimate(
  Y = pred_real_Ycons$Y,
  D = pred_real_Ycons$D,
  S_e = pred_real_Ycons$S_e,
  S_o = pred_real_Ycons$S_o,
  pred_Y = pred_real_Ycons$pred_Y,
  pred_D = pred_real_Ycons$pred_D,
  pred_S_e = pred_real_Ycons$pred_S_e,
  pred_S_o = pred_real_Ycons$pred_S_o,
  eps = 1e-2,
  method = "none",
  se = TRUE,
  se_params = list(B = 1000, fix_seed = TRUE, cores = cores, clusters = pred_real_Ycons$clusters),
)
print(rsv_real_Ycons)
```

```{r rsv-real-Ylowinc, message=FALSE}
# Low income prediction data (included in package)
# pred_real_Ylowinc
rsv_real_Ylowinc <- rsv_estimate(
  Y = pred_real_Ylowinc$Y,
  D = pred_real_Ylowinc$D,
  S_e = pred_real_Ylowinc$S_e,
  S_o = pred_real_Ylowinc$S_o,
  pred_Y = pred_real_Ylowinc$pred_Y,
  pred_D = pred_real_Ylowinc$pred_D,
  pred_S_e = pred_real_Ylowinc$pred_S_e,
  pred_S_o = pred_real_Ylowinc$pred_S_o,
  eps = 1e-2,
  method = "none",
  se = TRUE,
  se_params = list(B = 1000, fix_seed = TRUE, cores = cores, clusters = pred_real_Ylowinc$clusters),
)
print(rsv_real_Ylowinc)
```

```{r rsv-real-Ymidinc, message=FALSE}
# Middle income prediction data (included in package)
# pred_real_Ymidinc
rsv_real_Ymidinc <- rsv_estimate(
  Y = pred_real_Ymidinc$Y,
  D = pred_real_Ymidinc$D,
  S_e = pred_real_Ymidinc$S_e,
  S_o = pred_real_Ymidinc$S_o,
  pred_Y = pred_real_Ymidinc$pred_Y,
  pred_D = pred_real_Ymidinc$pred_D,
  pred_S_e = pred_real_Ymidinc$pred_S_e,
  pred_S_o = pred_real_Ymidinc$pred_S_o,
  eps = 1e-2,
  method = "none",
  se = TRUE,
  se_params = list(B = 1000, fix_seed = TRUE, cores = cores, clusters = pred_real_Ymidinc$clusters),
)
print(rsv_real_Ymidinc)
```

## RSV Estimation with Synthetic Sample Definition

We next consider a **synthetic sample definition** that differs from the real design used in the study.
In this synthetic specification:

- **Experimental sample** (\(S_e\)): Villages that were part of the randomized evaluation
  - Experimental: Treated (2010)
  - Experimental: Untreated (2011)
  - Experimental: Untreated (2012)
  
- **Observational sample** (\(S_o\)): Villages located within the first 50% of clusters (subdistricts).

```{r tab-synthS, echo = FALSE}
# Verify sample structure
data_synth %>%
  group_by(S) %>%
  summarise(
    `D = 0` = sum(D == 0, na.rm = TRUE),
    `D = 1` = sum(D == 1, na.rm = TRUE),
    `D = NA` = sum(is.na(D)),
    `Ycons = 0` = sum(Ycons == 0, na.rm = TRUE),
    `Ycons = 1` = sum(Ycons == 1, na.rm = TRUE),
    `Ycons = NA` = sum(is.na(Ycons)),
    `Ylowinc = 0` = sum(Ylowinc == 0, na.rm = TRUE),
    `Ylowinc = 1` = sum(Ylowinc == 1, na.rm = TRUE),
    `Ylowinc = NA` = sum(is.na(Ylowinc)),
    `Ymidinc = 0` = sum(Ymidinc == 0, na.rm = TRUE),
    `Ymidinc = 1` = sum(Ymidinc == 1, na.rm = TRUE),
    `Ymidinc = NA` = sum(is.na(Ymidinc))
  ) %>%
  pivot_longer(cols = -S, names_to = "Variable", values_to = "Count") %>%
  pivot_wider(names_from = S, values_from = Count) %>%
  kable(format = "markdown", digits = 3, caption = "Distribution of D and Y by S (Synthetic Definition)", col.names = c("Varibale & Value", "Experimental & Observational Sample", "Experimental Sample Only", "Observational Sample Only", "Neither (Excluded)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

We now estimate treatment effects for each poverty outcome under this synthetic sample definition.
The estimation procedure is identical to that in the real-sample case, except that here the sample indicators (`S_e`, `S_o`) are synthetically generated.

```{r rsv-synth-Ycons, message=FALSE}
# Consumption prediction data (included in package)
# pred_synth_Ycons
rsv_synth_Ycons <- rsv_estimate(
  Y = pred_synth_Ycons$Y,
  D = pred_synth_Ycons$D,
  S_e = pred_synth_Ycons$S_e,
  S_o = pred_synth_Ycons$S_o,
  pred_Y = pred_synth_Ycons$pred_Y,
  pred_D = pred_synth_Ycons$pred_D,
  pred_S_e = pred_synth_Ycons$pred_S_e,
  pred_S_o = pred_synth_Ycons$pred_S_o,
  eps = 1e-2,
  method = "none",
  se = TRUE,
  se_params = list(B = 1000, fix_seed = TRUE, cores = cores, clusters = pred_synth_Ycons$clusters),
)
print(rsv_synth_Ycons)
```

```{r rsv-synth-Ylowinc, message=FALSE}
# Low income prediction data (included in package)
# pred_synth_Ylowinc
rsv_synth_Ylowinc <- rsv_estimate(
  Y = pred_synth_Ylowinc$Y,
  D = pred_synth_Ylowinc$D,
  S_e = pred_synth_Ylowinc$S_e,
  S_o = pred_synth_Ylowinc$S_o,
  pred_Y = pred_synth_Ylowinc$pred_Y,
  pred_D = pred_synth_Ylowinc$pred_D,
  pred_S_e = pred_synth_Ylowinc$pred_S_e,
  pred_S_o = pred_synth_Ylowinc$pred_S_o,
  eps = 1e-2,
  method = "none",
  se = TRUE,
  se_params = list(B = 1000, fix_seed = TRUE, cores = cores, clusters = pred_synth_Ylowinc$clusters),
)
print(rsv_synth_Ylowinc)
```

```{r rsv-synth-Ymidinc, message=FALSE}
# Middle income prediction data (included in package)
# pred_synth_Ymidinc
rsv_synth_Ymidinc <- rsv_estimate(
  Y = pred_synth_Ymidinc$Y,
  D = pred_synth_Ymidinc$D,
  S_e = pred_synth_Ymidinc$S_e,
  S_o = pred_synth_Ymidinc$S_o,
  pred_Y = pred_synth_Ymidinc$pred_Y,
  pred_D = pred_synth_Ymidinc$pred_D,
  pred_S_e = pred_synth_Ymidinc$pred_S_e,
  pred_S_o = pred_synth_Ymidinc$pred_S_o,
  eps = 1e-2,
  method = "none",
  se = TRUE,
  se_params = list(B = 1000, fix_seed = TRUE, cores = cores, clusters = pred_synth_Ymidinc$clusters),
)
print(rsv_synth_Ymidinc)
```

## Benchmark Estimation

For comparison, we estimate the treatment effect using **only** the experimental sample--i.e., villages where the treatment indicator \(D\) and outcome \(Y\) are observed. This "experimental-only" benchmark ignores the observational sample and therefore does not leverage RSVs.

```{r benchmark-Ycons, message=FALSE}
# Filter to experimental sample with observed outcomes
data_exp_Ycons <- data %>%
  filter(
    !is.na(D),
    !is.na(Ycons)
    )

# Benchmark regression with cluster-robust SEs
benchmark_Ycons <- feols(Ycons ~ D, data = data_exp_Ycons, cluster = ~clusters)
summary(benchmark_Ycons)
```


```{r benchmark-Ylowinc, message=FALSE}
# Filter to experimental sample with observed outcomes
data_exp_Ylowinc <- data %>%
  filter(
    !is.na(D),
    !is.na(Ylowinc)
    )

# Benchmark regression with cluster-robust SEs
benchmark_Ylowinc <- feols(Ylowinc ~ D, data = data_exp_Ylowinc, cluster = ~clusters)
summary(benchmark_Ylowinc)
```


```{r benchmark-Ymidinc, message=FALSE}
# Filter to experimental sample with observed outcomes
data_exp_Ymidinc <- data %>%
  filter(
    !is.na(D),
    !is.na(Ymidinc)
    )

# Benchmark regression with cluster-robust SEs
benchmark_Ymidinc <- feols(Ymidinc ~ D, data = data_exp_Ymidinc, cluster = ~clusters)
summary(benchmark_Ymidinc)
```

# Treatment Effect Estimates

We summarize the estimated treatment effects and 90% confidence intervals for three estimators: a benchmark specification, the RSV estimator with a synthetic sample definition, and the RSV estimator with the real sample definition used in the paper.

```{r compile-results}
# Helper function to extract results
extract <- function(model) {
  coef_val <- coef(model)["D"]
  se_val <- model$se["D"]
  ci <- unname(confint(model, parm = "D", level = 0.90))

  data.frame(
    coef = coef_val,
    se = se_val,
    lci = ci[1],
    uci = ci[2],
    row.names = NULL
  )
}

results <- rbind(
  data.frame(outcome="Consumption", estimator = "Benchmark", extract(benchmark_Ycons)),
  data.frame(outcome="Consumption", estimator = "RSV: Synthetic samples", extract(rsv_synth_Ycons)),
  data.frame(outcome="Consumption", estimator = "RSV: Real samples", extract(rsv_real_Ycons)),
  data.frame(outcome="Low Income", estimator = "Benchmark", extract(benchmark_Ylowinc)),
  data.frame(outcome="Low Income", estimator = "RSV: Synthetic samples", extract(rsv_synth_Ylowinc)),
  data.frame(outcome="Low Income", estimator = "RSV: Real samples", extract(rsv_real_Ylowinc)),
  data.frame(outcome="Middle Income", estimator = "Benchmark", extract(benchmark_Ymidinc)),
  data.frame(outcome="Middle Income", estimator = "RSV: Synthetic samples", extract(rsv_synth_Ymidinc)),
  data.frame(outcome="Middle Income", estimator = "RSV: Real samples", extract(rsv_real_Ymidinc))
)
```


```{r results-table, echo=FALSE}
kable(
  results, 
  digits = 3, 
  row.names = FALSE,
  format = "markdown",
  align = "llrrrr",
  col.names = c("Outcome", "Estimator", "Coefficient", "SE", "Lower CI", "Upper CI"),
  caption = "Treatment Effects"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

We also visualize the point estimates with their 90% confidence intervals for each outcome.

```{r plot-function}
# Plot function
plot_treatment_effects <- function(results, title=waiver()) {
  results$estimator <- factor(results$estimator,levels = c(
    "Benchmark",
    "RSV: Synthetic samples",
    "RSV: Real samples"
  ))
  
  ggplot(results, aes(x = estimator, y = coef, shape = estimator, color = estimator)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
    geom_errorbar(aes(ymin=lci, ymax=uci), linewidth=0.65, width = 0.45) +
    geom_point(size = 2.5) +
    labs(
      title = title,
      x = "",
      y = TeX("Treatment effect $\\theta$"),
      shape=NULL,
      color=NULL
    ) +
    scale_color_manual(values = c(
      "Benchmark" = "black",
      "RSV: Synthetic samples" = "#8DB1CE",
      "RSV: Real samples" = "#43739D"
    )) +
    scale_shape_manual(values = c(
      "Benchmark" = 17,
      "RSV: Synthetic samples" = 15,
      "RSV: Real samples" = 19
    )) +
    scale_x_discrete(labels = ~ str_wrap(as.character(.x), 5, whitespace_only=F)) +
    scale_y_continuous(limits = c(-0.15, 0.01), minor_breaks = seq(-2,2,0.01)) +
    theme_bw() +
    theme(
      legend.position="none",
      panel.grid.major.x=element_blank(),
      panel.grid.minor.x=element_blank(),
      panel.grid.major.y=element_line(linewidth=0.3),
      panel.grid.minor.y=element_line(linewidth=0.1),
      text = element_text(size=14, family="Times", color="black"),
      axis.title.y = element_text(size=14, family="Times", color="black"),
      axis.title.x = element_blank(),
      axis.text = element_text(size=12, family="Times", color="black"),
    )
}
```

```{r plot-Ycons, fig.width=3, fig.height=3}
results %>% 
  filter(outcome=="Consumption") %>%
  plot_treatment_effects(title="Consumption")
```

```{r plot-Ylowinc, fig.width=3, fig.height=3}
results %>% 
  filter(outcome=="Low Income") %>%
  plot_treatment_effects(title="Low Income")
```

```{r plot-Ymidinc, fig.width=3, fig.height=3}
results %>% 
  filter(outcome=="Middle Income") %>%
  plot_treatment_effects(title="Middle Income")
```


<!-- # Denominator (Relevance) -->

<!-- The RSV estimator's denominator (relevance) summarizes how informative the remotely sensed variables are for predicting the outcome. We report the point estimate, standard error, and the 90% CI. -->

<!-- ```{r compile-denominator-results} -->
<!-- # Helper function to extract results -->
<!-- extract_denominator <- function(model) { -->
<!--   level = 0.90 -->
<!--   alpha = 1 - level -->
<!--   coef_val <- model$denominator -->
<!--   se_val <- model$denominator_se -->
<!--   ci <- c( -->
<!--     coef_val - qnorm(1 - alpha/2) * se_val,  -->
<!--     coef_val + qnorm(1 - alpha/2) * se_val -->
<!--   ) -->

<!--   data.frame( -->
<!--     coef = coef_val, -->
<!--     se = se_val, -->
<!--     lci = ci[1], -->
<!--     uci = ci[2], -->
<!--     row.names = NULL -->
<!--   ) -->
<!-- } -->

<!-- results_denominator <- rbind( -->
<!--   data.frame(outcome = "Consumption", estimator = "RSV: Synthetic samples", extract_denominator(rsv_synth_Ycons)), -->
<!--   data.frame(outcome = "Consumption", estimator = "RSV: Real samples", extract_denominator(rsv_real_Ycons)), -->
<!--   data.frame(outcome = "Low Income", estimator = "RSV: Synthetic samples", extract_denominator(rsv_synth_Ylowinc)), -->
<!--   data.frame(outcome = "Low Income", estimator = "RSV: Real samples", extract_denominator(rsv_real_Ylowinc)), -->
<!--   data.frame(outcome = "Middle Income", estimator = "RSV: Synthetic samples", extract_denominator(rsv_synth_Ymidinc)), -->
<!--   data.frame(outcome = "Middle Income", estimator = "RSV: Real samples", extract_denominator(rsv_real_Ymidinc)) -->
<!-- ) -->
<!-- ``` -->


<!-- ```{r results-denominator-table, echo = FALSE} -->
<!-- # Consumption poverty -->
<!-- kable( -->
<!--   results_denominator,  -->
<!--   digits = 3,  -->
<!--   row.names = FALSE, -->
<!--   format = "markdown", -->
<!--   align = "llrrrr", -->
<!--   col.names = c("Outcome", "Estimator", "Coefficient", "SE", "Lower CI", "Upper CI"), -->
<!--   caption = "Denominator (Relevance)" -->
<!--   ) %>% -->
<!--   kable_styling(bootstrap_options = c("striped", "hover")) -->
<!-- ``` -->


<!-- We also visualize the denominator estimates with their 90% confidence intervals for the synthetic and real sample definitions. -->

<!-- ```{r plot-denominator-function} -->
<!-- # Plot function -->
<!-- plot_denominator <- function(results, title=waiver()) { -->
<!--   results$estimator <- factor(results$estimator,levels = c( -->
<!--     "RSV: Synthetic samples", -->
<!--     "RSV: Real samples" -->
<!--   )) -->

<!--   ggplot(results, aes(x = estimator, y = coef, shape = estimator, color = estimator)) + -->
<!--     geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") + -->
<!--     geom_errorbar(aes(ymin=lci, ymax=uci), linewidth=0.65, width = 0.28) + -->
<!--     geom_point(size = 2.5) + -->
<!--     labs( -->
<!--       title = title, -->
<!--       x = "", -->
<!--       y = "Denominator (Relevance)", -->
<!--       shape=NULL, -->
<!--       color=NULL -->
<!--     ) + -->
<!--     scale_color_manual(values = c( -->
<!--       "RSV: Synthetic samples" = "#8DB1CE", -->
<!--       "RSV: Real samples" = "#43739D" -->
<!--     )) + -->
<!--     scale_shape_manual(values = c( -->
<!--       "RSV: Synthetic samples" = 15, -->
<!--       "RSV: Real samples" = 19 -->
<!--     )) + -->
<!--     scale_x_discrete(labels = ~ str_wrap(as.character(.x), 5, whitespace_only=F)) + -->
<!--     scale_y_continuous(limits = c(0,6.1), minor_breaks = seq(-2,10,0.25)) + -->
<!--     theme_bw() + -->
<!--     theme( -->
<!--       legend.position="none", -->
<!--       panel.grid.major.x=element_blank(), -->
<!--       panel.grid.minor.x=element_blank(), -->
<!--       panel.grid.major.y=element_line(linewidth=0.3), -->
<!--       panel.grid.minor.y=element_line(linewidth=0.1), -->
<!--       text = element_text(size=14, family="Times", color="black"), -->
<!--       axis.title.y = element_text(size=14, family="Times", color="black"), -->
<!--       axis.title.x = element_blank(), -->
<!--       axis.text = element_text(size=12, family="Times", color="black"), -->
<!--     ) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r plot-denominator-Ycons, fig.width=3, fig.height=3} -->
<!-- results_denominator %>%  -->
<!--   filter(outcome=="Consumption") %>% -->
<!--   plot_denominator(title="Consumption") -->
<!-- ``` -->

<!-- ```{r plot-denominator-Ylowinc, fig.width=3, fig.height=3} -->
<!-- results_denominator %>%  -->
<!--   filter(outcome=="Low Income") %>% -->
<!--   plot_denominator(title="Low Income") -->
<!-- ``` -->

<!-- ```{r plot-denominator-Ymidinc, fig.width=3, fig.height=3} -->
<!-- results_denominator %>%  -->
<!--   filter(outcome=="Middle Income") %>% -->
<!--   plot_denominator(title="Middle Income") -->
<!-- ``` -->


# References

Muralidharan, K., Niehaus, P., and Sukhtankar, S. (2017). "General Equilibrium Effects of (Improving) Public Employment Programs: Experimental Evidence from India." *NBER Working Paper 23838*.

Rambachan, A., Singh, R., and Viviano, D. (2025). "Program Evaluation with Remotely Sensed Outcomes." *arXiv:2411.10959*.

# Session Info

```{r session-info}
sessionInfo()
```
