---
title: "Estimating Treatment Effects with Remotely Sensed Variables"
author: "Rambachan, Singh, and Viviano (2025)"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Estimating Treatment Effects with Remotely Sensed Variables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
  #, eval = FALSE
)
```

# Overview

This vignette demonstrates how to use the `remoteoutcome` package to estimate treatment effects when outcomes are measured using remotely sensed variables (RSVs). It illustrates how **Figure 7**  of *Rambachan, Singh, and Viviano (2025)* could be produced using `remoteoutcome`, which analyzes data from a randomized evaluation of an anti-poverty program in Andhra Pradesh, India (Muralidharan et al., 2016; Muralidharan et al. 2023). We specifically illustrate how `remoteoutcome` can be used to (i) 
estimate predictions of outcomes, treatment, and the sample indicator using remotely sensed variables, and (ii) report estimates of the treatment effect using cross-fitting.

# Data

The dataset combines experimental and remotely sensed information used in the poverty application.  
It includes:

- **Outcome variables:** Poverty indicators such as household consumption and income measures.  
- **Treatment variable:** Program assignment indicating whether a village received the Smartcard intervention.  
- **Remotely sensed variables:** Satellite-based measures of luminosity and related features.  
- **Sample indicators:** Experimental design identifiers distinguishing treatment and control groups.

```{r load-data}
library(latex2exp)
library(stringr)
library(knitr)
library(tibble)
library(kableExtra)
library(dplyr, warn.conflicts = F)
library(ggplot2, warn.conflicts = F)
library(fixest)
library(tidyr)
library(remoteoutcome)

cores <- parallel::detectCores() - 1
```

```{r tab-data, echo=FALSE}
# Load the data
data(smartcard_data, library="remoteoutcome")
data(remote_vars_p1, library="remoteoutcome")
data(remote_vars_p2, library="remoteoutcome")
data(remote_vars_p3, library="remoteoutcome")

# Merge remote variables
smartcard_data <- smartcard_data %>%
  inner_join(remote_vars_p1, by="shrid2") %>%
  inner_join(remote_vars_p2, by="shrid2") %>%
  inner_join(remote_vars_p3, by="shrid2")

smartcard_data %>% 
  select(shrid2, clusters, `Sample (Smartcard)`, D, Ylowinc, Ymidinc, Ycons, luminosity_max.2012, luminosity_max.2013, satellite_1, satellite_2) %>%
  group_by(D) %>%
  slice(1:2) %>%
  ungroup() %>%
  head()  %>%
  kable(digits = 3, caption="Data overview") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r}
cat(sprintf("Total observations: %d\n", nrow(smartcard_data)))
cat(sprintf("Number of clusters (subdistricts): %d\n",
            length(unique(smartcard_data$clusters))))
```

The poverty outcomes analyzed in the application are defined as follows:

  - `Ycons`: Household consumption below 25th percentile.
  - `Ylowinc`: No household members earning more than 5,000 INR per month.
  - `Ymidinc`: No household members earning more than 10,000 INR per month.
  
```{r tab-outcomes, echo=FALSE}
smartcard_data %>%
  select(Ycons, Ylowinc, Ymidinc) %>%
  rename(
    "Consumption" = Ycons,
    "Low Income" = Ylowinc,
    "Middle Income" = Ymidinc
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Outcome",
    values_to = "value"
  ) %>%
  group_by(Outcome) %>%
  summarise(
    n = sum(!is.na(value)),
    mean = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    min = min(value, na.rm = TRUE),
    max = max(value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  kable(digits = 3, caption = "Outcome Variables") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

The experiment was implemented in multiple phases, with villages randomly assigned to receive Smartcards in different years:

- **Experimental: Treated (2010)**: Villages that received Smartcards in 2010
- **Experimental: Untreated (2011)**: Buffer villages assigned to receive Smartcards in 2011.
- **Experimental: Untreated (2012)**: Control villages assigned to receive Smartcards in 2012.
- **Observational (N/A)**: Non-study or holdout villages excluded from the experiment.

```{r tab-exp-design, echo=FALSE}
tab <- smartcard_data %>% 
  group_by(`Sample (Smartcard)`) %>%
  summarise(
    `Number of villages`        = n_distinct(shrid2),
    `Average population`        = round(mean(tot_p), 0),
    `Average fraction female`   = round(mean(tot_f), 3),
    .groups = "drop"
  ) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column(var = "Metric")

## Use first row as column names
colnames(tab) <- tab[1, ]
tab <- tab[-1, ]

## Create markdown table
kable(tab, row.names = FALSE, caption="Experimental Design") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Analysis

## RSV Estimation with Real Sample Definition

We demonstrates how to estimate treatment effects using the **RSV estimator** under the *real sample definition*.

Following the study design, we construct two key sample indicators:

- **Experimental sample** (\(S_e\)): Villages that were part of the randomized evaluation
  - Experimental: Treated (2010)
  - Experimental: Untreated (2011)
  - Experimental: Untreated (2012)
  
- **Observational sample** (\(S_o\)): 
  - Experimental: Untreated (2011)
  - Observational (N/A)

This creates an overlap where *Experimental: Untreated (2011)* villages are in both samples.

```{r real-sample-def}
data_real <- create_data_real(smartcard_data)
```

```{r tab-realS, echo = FALSE}
# Verify sample structure
data_real %>%
  group_by(S) %>%
  summarise(
    `D = 0` = sum(D == 0, na.rm = TRUE),
    `D = 1` = sum(D == 1, na.rm = TRUE),
    `D = NA` = sum(is.na(D)),
    `Ycons = 0` = sum(Ycons == 0, na.rm = TRUE),
    `Ycons = 1` = sum(Ycons == 1, na.rm = TRUE),
    `Ycons = NA` = sum(is.na(Ycons)),
    `Ylowinc = 0` = sum(Ylowinc == 0, na.rm = TRUE),
    `Ylowinc = 1` = sum(Ylowinc == 1, na.rm = TRUE),
    `Ylowinc = NA` = sum(is.na(Ylowinc)),
    `Ymidinc = 0` = sum(Ymidinc == 0, na.rm = TRUE),
    `Ymidinc = 1` = sum(Ymidinc == 1, na.rm = TRUE),
    `Ymidinc = NA` = sum(is.na(Ymidinc))
  ) %>%
  pivot_longer(cols = -S, names_to = "Variable", values_to = "Count") %>%
  pivot_wider(names_from = S, values_from = Count) %>%
  kable(format = "markdown", digits = 3, caption = "Distribution of D and Y by S (Real Definition)", col.names = c("Varibale & Value", "Experimental & Observational Sample", "Experimental Sample Only", "Observational Sample Only")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

We now estimate treatment effects for each poverty outcome using the `rsv_estimate()` function.
This function implements the RSV estimator described in **Rambachan, Singh, and Viviano (2025)**.

In this vignette, we train predictive models using `ranger` library. 

Key arguments in this example:

- `Y`, `D`, `S_e`, `S_o`: True observed values of outcomes, treatment assignment, and sample indicators.
- `R`: Matrix of remotely sensed predictors, including luminosity and satellite-derived features.
- `method = "none"`: Uses all available data for both training and estimation (no cross-fitting).
- `ml_params`: Random forest parameters:
  - `ntree = 100`: Number of trees.
  - `classwt_Y`: Weights for classes `Y=0` and `Y=1` for the \(\text{PRED}_Y(R)\) model.
  - `seed`: User specified seed passed to each `ranger` function for reproducibility
- `se = TRUE`: Requests bootstrap-based standard errors with cluster resampling.
- `se_params`: Controls bootstrap settings:
  - `B = 1000`: Number of bootstrap replications.
  - `clusters`: Subdistrict-level cluster identifiers.
  - `fix_seed = TRUE`: Ensures deterministic seeding for reproducibility.
- `cores`: Number of cores used by either `ranger` or bootstrap computation.

```{r rsv-real-Ycons, message=FALSE}
# Consumption poverty
Y <- data_real$Ycons
D <- data_real$D
R <- data_real %>% select(starts_with("luminosity"), starts_with("satellite"))
S_e <- !is.na(D) & (rowSums(is.na(R)) == 0)
S_o <- !is.na(Y) & (rowSums(is.na(R)) == 0)
clusters <- data_real$clusters

rsv_real_Ycons <- rsv_estimate(
  Y = Y,
  D = D,
  R = R,
  S_e = S_e,
  S_o = S_o,
  eps = 1e-2,
  method = "none",
  ml_params = list(ntree = 100, classwt_Y = c(10, 1), seed=42),
  se = TRUE,
  se_params = list(B = 1000, fix_seed = TRUE, clusters = clusters),
  cores = cores
)
print(rsv_real_Ycons)
```

```{r rsv-real-Ylowinc, message=FALSE}
# Low income poverty
Y <- data_real$Ylowinc
D <- data_real$D
R <- data_real %>% select(starts_with("luminosity"), starts_with("satellite"))
S_e <- !is.na(D) & (rowSums(is.na(R)) == 0)
S_o <- !is.na(Y) & (rowSums(is.na(R)) == 0)
clusters <- data_real$clusters

rsv_real_Ylowinc <- rsv_estimate(
  Y = Y,
  D = D,
  R = R,
  S_e = S_e,
  S_o = S_o,
  eps = 1e-2,
  method = "none",
  ml_params = list(ntree = 100, classwt_Y = c(10, 1), seed=42),
  se = TRUE,
  se_params = list(B = 1000, fix_seed = TRUE, clusters = clusters),
  cores = cores
)
print(rsv_real_Ylowinc)
```

```{r rsv-real-Ymidinc, message=FALSE}
# Middle income poverty
Y <- data_real$Ymidinc
D <- data_real$D
R <- data_real %>% select(starts_with("luminosity"), starts_with("satellite"))
S_e <- !is.na(D) & (rowSums(is.na(R)) == 0)
S_o <- !is.na(Y) & (rowSums(is.na(R)) == 0)
clusters <- data_real$clusters

rsv_real_Ymidinc <- rsv_estimate(
  Y = Y,
  D = D,
  R = R,
  S_e = S_e,
  S_o = S_o,
  eps = 1e-2,
  method = "none",
  ml_params = list(ntree = 100, classwt_Y = c(10, 1), seed=42),
  se = TRUE,
  se_params = list(B = 1000, fix_seed = TRUE, clusters = clusters),
  cores = cores
)
print(rsv_real_Ymidinc)
```


## RSV Estimation with Synthetic Sample Definition

We next consider a **synthetic sample definition** that differs from the real design used in the study.
In this synthetic specification:

- **Experimental sample** (\(S_e\)): Villages that were part of the randomized evaluation
  - Experimental: Treated (2010)
  - Experimental: Untreated (2011)
  - Experimental: Untreated (2012)
  
- **Observational sample** (\(S_o\)): Villages located within the first 50% of clusters (subdistricts).

```{r synth-sample-def}
data_synth <- create_data_synth(smartcard_data)
```

```{r tab-synthS, echo = FALSE}
# Verify sample structure
data_synth %>%
  group_by(S) %>%
  summarise(
    `D = 0` = sum(D == 0, na.rm = TRUE),
    `D = 1` = sum(D == 1, na.rm = TRUE),
    `D = NA` = sum(is.na(D)),
    `Ycons = 0` = sum(Ycons == 0, na.rm = TRUE),
    `Ycons = 1` = sum(Ycons == 1, na.rm = TRUE),
    `Ycons = NA` = sum(is.na(Ycons)),
    `Ylowinc = 0` = sum(Ylowinc == 0, na.rm = TRUE),
    `Ylowinc = 1` = sum(Ylowinc == 1, na.rm = TRUE),
    `Ylowinc = NA` = sum(is.na(Ylowinc)),
    `Ymidinc = 0` = sum(Ymidinc == 0, na.rm = TRUE),
    `Ymidinc = 1` = sum(Ymidinc == 1, na.rm = TRUE),
    `Ymidinc = NA` = sum(is.na(Ymidinc))
  ) %>%
  pivot_longer(cols = -S, names_to = "Variable", values_to = "Count") %>%
  pivot_wider(names_from = S, values_from = Count) %>%
  kable(format = "markdown", digits = 3, caption = "Distribution of D and Y by S (Synthetic Definition)", col.names = c("Varibale & Value", "Experimental & Observational Sample", "Experimental Sample Only", "Observational Sample Only", "Neither (Excluded)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

We now estimate treatment effects for each poverty outcome under this synthetic sample definition.
The estimation procedure is identical to that in the real-sample case, except that here the sample indicators (`S_e`, `S_o`) are synthetically generated.

```{r rsv-synth, message=FALSE}
# Consumption poverty
Y <- data_synth$Ycons
D <- data_synth$D
R <- data_synth %>% select(starts_with("luminosity"), starts_with("satellite"))
S_e <- !is.na(D) & (rowSums(is.na(R)) == 0)
S_o <- !is.na(Y) & (rowSums(is.na(R)) == 0)
clusters <- data_synth$clusters

rsv_synth_Ycons <- rsv_estimate(
  Y = Y,
  D = D,
  R = R,
  S_e = S_e,
  S_o = S_o,
  eps = 1e-2,
  method = "none",
  ml_params = list(ntree = 100, classwt_Y = c(10, 1), seed=42),
  se = TRUE,
  se_params = list(B = 1000, fix_seed = TRUE, clusters = clusters),
  cores = cores
)
print(rsv_synth_Ycons)
```

```{r rsv-synth-Ylowinc, message=FALSE}
# Low income poverty
Y <- data_synth$Ylowinc
D <- data_synth$D
R <- data_synth %>% select(starts_with("luminosity"), starts_with("satellite"))
S_e <- !is.na(D) & (rowSums(is.na(R)) == 0)
S_o <- !is.na(Y) & (rowSums(is.na(R)) == 0)
clusters <- data_synth$clusters

rsv_synth_Ylowinc <- rsv_estimate(
  Y = Y,
  D = D,
  R = R,
  S_e = S_e,
  S_o = S_o,
  eps = 1e-2,
  method = "none",
  ml_params = list(ntree = 100, classwt_Y = c(10, 1), seed=42),
  se = TRUE,
  se_params = list(B = 1000, fix_seed = TRUE, clusters = clusters),
  cores = cores
)
print(rsv_synth_Ylowinc)
```

```{r rsv-synth-Ymidinc, message=FALSE}
# Middle income poverty
Y <- data_synth$Ymidinc
D <- data_synth$D
R <- data_synth %>% select(starts_with("luminosity"), starts_with("satellite"))
S_e <- !is.na(D) & (rowSums(is.na(R)) == 0)
S_o <- !is.na(Y) & (rowSums(is.na(R)) == 0)
clusters <- data_synth$clusters

rsv_synth_Ymidinc <- rsv_estimate(
  Y = Y,
  D = D,
  R = R,
  S_e = S_e,
  S_o = S_o,
  eps = 1e-2,
  method = "none",
  ml_params = list(ntree = 100, classwt_Y = c(10, 1), seed=42),
  se = TRUE,
  se_params = list(B = 1000, fix_seed = TRUE, clusters = clusters),
  cores = cores
)
print(rsv_synth_Ymidinc)
```


## Benchmark Estimation

For comparison, we estimate the treatment effect using **only** the experimental sample--i.e., villages where the treatment indicator \(D\) and outcome \(Y\) are observed. This "experimental-only" benchmark ignores the observational sample and therefore does not leverage RSVs.

```{r benchmark-Ycons, message=FALSE}
# Filter to experimental sample with observed outcomes
data_exp_Ycons <- smartcard_data %>%
  filter(
    !is.na(D),
    !is.na(Ycons)
    )

# Benchmark regression with cluster-robust SEs
benchmark_Ycons <- feols(Ycons ~ D, data = data_exp_Ycons, cluster = ~clusters)
summary(benchmark_Ycons)
```


```{r benchmark-Ylowinc, message=FALSE}
# Filter to experimental sample with observed outcomes
data_exp_Ylowinc <- smartcard_data %>%
  filter(
    !is.na(D),
    !is.na(Ylowinc)
    )

# Benchmark regression with cluster-robust SEs
benchmark_Ylowinc <- feols(Ylowinc ~ D, data = data_exp_Ylowinc, cluster = ~clusters)
summary(benchmark_Ylowinc)
```


```{r benchmark-Ymidinc, message=FALSE}
# Filter to experimental sample with observed outcomes
data_exp_Ymidinc <- smartcard_data %>%
  filter(
    !is.na(D),
    !is.na(Ymidinc)
    )

# Benchmark regression with cluster-robust SEs
benchmark_Ymidinc <- feols(Ymidinc ~ D, data = data_exp_Ymidinc, cluster = ~clusters)
summary(benchmark_Ymidinc)
```


# Treatment Effect Estimates

We summarize the estimated treatment effects and 90% confidence intervals for three estimators: a benchmark specification, the RSV estimator with a synthetic sample definition, and the RSV estimator with the real sample definition used in the paper.

```{r compile-results}
# Helper function to extract results
extract <- function(model) {
  coef_val <- coef(model)["D"]
  se_val <- model$se["D"]
  ci <- unname(confint(model, parm = "D", level = 0.90))

  data.frame(
    coef = coef_val,
    se = se_val,
    lci = ci[1],
    uci = ci[2],
    row.names = NULL
  )
}

results <- rbind(
  data.frame(outcome="Consumption", estimator = "Benchmark", extract(benchmark_Ycons)),
  data.frame(outcome="Consumption", estimator = "RSV: Synthetic samples", extract(rsv_synth_Ycons)),
  data.frame(outcome="Consumption", estimator = "RSV: Real samples", extract(rsv_real_Ycons)),
  data.frame(outcome="Low Income", estimator = "Benchmark", extract(benchmark_Ylowinc)),
  data.frame(outcome="Low Income", estimator = "RSV: Synthetic samples", extract(rsv_synth_Ylowinc)),
  data.frame(outcome="Low Income", estimator = "RSV: Real samples", extract(rsv_real_Ylowinc)),
  data.frame(outcome="Middle Income", estimator = "Benchmark", extract(benchmark_Ymidinc)),
  data.frame(outcome="Middle Income", estimator = "RSV: Synthetic samples", extract(rsv_synth_Ymidinc)),
  data.frame(outcome="Middle Income", estimator = "RSV: Real samples", extract(rsv_real_Ymidinc))
)
```


```{r results-table, echo=FALSE}
kable(
  results, 
  digits = 3, 
  row.names = FALSE,
  format = "markdown",
  align = "llrrrr",
  col.names = c("Outcome", "Estimator", "Coefficient", "SE", "Lower CI", "Upper CI"),
  caption = "Treatment Effects"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

We also visualize the point estimates with their 90% confidence intervals for each outcome.

```{r plot-function}
# Plot function
plot_treatment_effects <- function(results, title=waiver()) {
  results$estimator <- factor(results$estimator,levels = c(
    "Benchmark",
    "RSV: Synthetic samples",
    "RSV: Real samples"
  ))
  
  ggplot(results, aes(x = estimator, y = coef, shape = estimator, color = estimator)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
    geom_errorbar(aes(ymin=lci, ymax=uci), linewidth=0.65, width = 0.45) +
    geom_point(size = 2.5) +
    labs(
      title = title,
      x = "",
      y = TeX("Treatment effect $\\theta$"),
      shape=NULL,
      color=NULL
    ) +
    scale_color_manual(values = c(
      "Benchmark" = "black",
      "RSV: Synthetic samples" = "#8DB1CE",
      "RSV: Real samples" = "#43739D"
    )) +
    scale_shape_manual(values = c(
      "Benchmark" = 17,
      "RSV: Synthetic samples" = 15,
      "RSV: Real samples" = 19
    )) +
    scale_x_discrete(labels = ~ str_wrap(as.character(.x), 5, whitespace_only=F)) +
    scale_y_continuous(limits = c(-0.15, 0.01), minor_breaks = seq(-2,2,0.01)) +
    theme_bw() +
    theme(
      legend.position="none",
      panel.grid.major.x=element_blank(),
      panel.grid.minor.x=element_blank(),
      panel.grid.major.y=element_line(linewidth=0.3),
      panel.grid.minor.y=element_line(linewidth=0.1),
      text = element_text(size=14, family="Times", color="black"),
      axis.title.y = element_text(size=14, family="Times", color="black"),
      axis.title.x = element_blank(),
      axis.text = element_text(size=12, family="Times", color="black"),
    )
}
```

```{r plot-Ycons, fig.width=3, fig.height=3}
results %>% 
  filter(outcome=="Consumption") %>%
  plot_treatment_effects(title="Consumption")
```

```{r plot-Ylowinc, fig.width=3, fig.height=3}
results %>% 
  filter(outcome=="Low Income") %>%
  plot_treatment_effects(title="Low Income")
```

```{r plot-Ymidinc, fig.width=3, fig.height=3}
results %>% 
  filter(outcome=="Middle Income") %>%
  plot_treatment_effects(title="Middle Income")
```

# Additional Notes

This vignette introduces two implementation changes relative to *Rambachan, Singh, and Viviano (2025)*:

1. The paper estimated \(\text{PRED}_Y(R), \text{PRED}_D(R), \text{PRED}_{S_e}(R), \text{PRED}_{S_o}(R)\) using `randomForest` package This vignette instead uses `ranger`, a modern random forest package, that improves speed and reproducibility.

2. The paper computed the bootstrap standard errors using the `boot` package with parallelization without setting the seed. `remoteoutcome` uses a custom cluster bootstrap that supports deterministic seeding via `fix_seed = TRUE`, setting `set.seed(b)` for the *b*‑th replication.

Under these changes, the vignette reproduces the point estimates as reported in the paper and there are small numerical differences in the endpoints of the confidence intervals.

# References

Muralidharan, K., Niehaus, P., and Sukhtankar, S. (2016). "Building state capacity: Evidence from biometric smartcards in India." *American Economic Review* 106 (10), 2895-2929.

Muralidharan, K., Niehaus, P., and Sukhtankar, S. (2023). "General equilibrium effects of (improving) public employment programs: Experimental evidence from India." *Econometrica* 91 (4), 1261-1295.

Rambachan, A., Singh, R., and Viviano, D. (2025). "Program Evaluation with Remotely Sensed Outcomes." *arXiv:2411.10959*.
