---
title: "Replication: Uganda Forest Cover"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replication: Uganda Forest Cover}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

run_chunks <- tryCatch({
  base_url <- paste0(
    "https://raw.githubusercontent.com/",
    "asheshrambachan/rsv_replication/main/data/clean/ugandaforestcover/"
  )
  con <- url(paste0(base_url, "data_Ybin.csv"))
  close(con)
  TRUE
}, error = function(e) FALSE, warning = function(w) FALSE)
```

## Overview

This vignette replicates the Uganda forest cover simulation study from
Rambachan, Singh, and Viviano using the **remoteoutcome** package.
The application studies the effect of a Payment for Ecosystem Services (PES)
programme on binary forest cover (`Y = 1` if cover exceeds the 80th percentile
of the imagery region).

The study design uses spatial proximity as the source of sample variation:

- **Experimental sample** (`S_e = 1`): 1,009 plots inside the PES imagery
  boundary — treatment `D` is observed, `Y` is also observed from satellite data.
- **Observational sample** (`S_o = 1`): 272 plots in the 0–2 km buffer just
  outside the boundary — `Y` is observed, `D` is not.
- **Remote sensing variable**: `R1 = P(Y=1|X)` — the random-forest predicted
  probability of high forest cover, where X are MOSAIKS satellite features.
  (`R0 + R1 = 1` by construction, so only `R1` is used to avoid perfect
  multicollinearity in the nuisance logits.)

The vignette covers three analyses:

1. **Stability test** — KS test comparing F(R1 | S = experimental, Y = y)
   against F(R1 | S = observational, Y = y).
2. **No experimental outcomes simulation** — repeated `cv.rsv()` on datasets
   where Y is never observed in the experimental sample.
3. **With experimental outcomes simulation** — same, but a subset of n_eo
   experimental units also have Y observed (overlap units, `S_e = S_o = 1`).

> **Note on simulation draws**: Both simulations use `B = 50` Monte Carlo
> replications with analytical influence-function SEs (no bootstrap loop),
> running in roughly 1 minute total. The published results use `B = 500` and
> score bootstrap (`B_se = 100`). With `B = 50`, coverage SEs are ≈ ±0.03;
> average SEs match the published values to within 0.001.
> See Sections 4 and 6 for side-by-side comparisons with the published numbers.

## Setup

```{r libraries, eval = run_chunks}
library(remoteoutcome)
```

The simulation DGP requires helper functions drawn from the replication
package utilities, defined locally here.

```{r helpers, eval = run_chunks}
# KL-tilting: find p1 s.t. E_{p1}[y] = E_{p0}[y] + tau
compute_p1_kl <- function(p0, y, tau, tol = 1e-10) {
  mu_target <- sum(y * p0) + tau
  support <- p0 > 0
  p0s <- p0[support]; ys <- y[support]
  if (abs(tau) < tol) return(p0)
  f <- function(lam) {
    z <- lam * ys - max(lam * ys)
    w <- p0s * exp(z); w <- w / sum(w)
    sum(ys * w) - mu_target
  }
  lam <- uniroot(f, c(-100, 100), tol = tol)$root
  z <- lam * ys - max(lam * ys); w <- p0s * exp(z); w <- w / sum(w)
  p1 <- numeric(length(p0)); p1[support] <- w; p1
}

# Draw Y for experimental units given treatment arm
sample_Y_exp <- function(D, y_levels, p0, p1) {
  K <- length(y_levels); out <- integer(length(D))
  idx0 <- D == 0; idx1 <- D == 1
  if (any(idx0)) out[idx0] <- sample.int(K, sum(idx0), replace = TRUE, prob = p0)
  if (any(idx1)) out[idx1] <- sample.int(K, sum(idx1), replace = TRUE, prob = p1)
  y_levels[out]
}

# Draw RSV rows from the empirical R|Y lookup table
sample_R_given_Y <- function(R_given_Y, y_vals, y_levels) {
  out <- matrix(NA_real_, length(y_vals), ncol(R_given_Y[[1]]))
  for (k in seq_along(y_levels)) {
    idx <- which(y_vals == y_levels[k])
    if (!length(idx)) next
    draw <- sample.int(nrow(R_given_Y[[k]]), length(idx), replace = TRUE)
    out[idx, ] <- R_given_Y[[k]][draw, , drop = FALSE]
  }
  out
}

# Generate one dataset: no experimental outcomes (S_e and S_o non-overlapping)
sim_noexp <- function(n_e, n_o, y_levels, p0, p1, alpha_o, R_given_Y) {
  D_e <- rbinom(n_e, 1, 0.5)
  Y_e <- sample_Y_exp(D_e, y_levels, p0, p1)
  R_e <- sample_R_given_Y(R_given_Y, Y_e, y_levels)
  Y_e <- rep(NA_real_, n_e)   # mask: Y unobserved in experimental sample
  K   <- length(y_levels)
  Y_o <- y_levels[sample.int(K, n_o, replace = TRUE, prob = alpha_o)]
  R_o <- sample_R_given_Y(R_given_Y, Y_o, y_levels)
  list(Y = c(Y_e, Y_o), D = c(D_e, rep(NA_real_, n_o)),
       R = rbind(R_e, R_o),
       S_e = c(rep(1L, n_e), rep(0L, n_o)),
       S_o = c(rep(0L, n_e), rep(1L, n_o)))
}

# Generate one dataset: with n_eo overlap units (S_e = S_o = 1)
sim_exp <- function(n_e, n_o, n_eo, y_levels, p0, p1, alpha_o, R_given_Y) {
  D_e <- rbinom(n_e, 1, 0.5)
  Y_e <- sample_Y_exp(D_e, y_levels, p0, p1)
  R_e <- sample_R_given_Y(R_given_Y, Y_e, y_levels)
  val_idx    <- sample.int(n_e, n_eo, replace = FALSE)
  pure_idx   <- setdiff(seq_len(n_e), val_idx)
  K          <- length(y_levels)
  Y_o <- y_levels[sample.int(K, n_o, replace = TRUE, prob = alpha_o)]
  R_o <- sample_R_given_Y(R_given_Y, Y_o, y_levels)
  list(
    Y   = c(Y_e[val_idx], rep(NA_real_, length(pure_idx)), Y_o),
    D   = c(D_e[val_idx], D_e[pure_idx],                  rep(NA_real_, n_o)),
    R   = rbind(R_e[val_idx, , drop=FALSE],
                R_e[pure_idx, , drop=FALSE], R_o),
    S_e = c(rep(1L, n_eo), rep(1L, length(pure_idx)), rep(0L, n_o)),
    S_o = c(rep(1L, n_eo), rep(0L, length(pure_idx)), rep(1L, n_o))
  )
}

# Summarise one tau block of simulation draws
summarise_sims <- function(draws, tau) {
  ci_mult <- qnorm(0.975)
  bias <- mean(draws$tauhat - tau, na.rm = TRUE)
  sd_e <- sd(draws$tauhat, na.rm = TRUE)
  cov  <- mean(tau >= draws$tauhat - ci_mult * draws$se &
               tau <= draws$tauhat + ci_mult * draws$se, na.rm = TRUE)
  data.frame(tau             = tau,
             normalized_bias = round(bias / sd_e, 3),
             coverage        = round(cov, 3),
             rmse            = round(sqrt(mean((draws$tauhat-tau)^2,na.rm=TRUE)), 4),
             avg_se          = round(mean(draws$se, na.rm = TRUE), 4),
             n_sims          = sum(!is.na(draws$tauhat)))
}
```

## 1. Load data

```{r load-data, eval = run_chunks}
base_url <- paste0(
  "https://raw.githubusercontent.com/",
  "asheshrambachan/rsv_replication/main/data/clean/ugandaforestcover/"
)
dt <- read.csv(paste0(base_url, "data_Ybin.csv"))

cat("Experimental (S_e=1):        ", sum(dt$S_e),      "\n")
cat("Observational 0-2km (S_o=1): ", sum(dt$S_o_02km), "\n")
cat("Y values:                    ", paste(sort(unique(dt$Y)), collapse=", "), "\n")
```

## 2. Stability assumption test

`test_stability()` compares F(R1 | S = experimental, Y = y) against
F(R1 | S = observational, Y = y) via a KS test for each outcome level.

```{r stability, eval = run_chunks}
R_score <- matrix(dt$R1, ncol = 1, dimnames = list(NULL, "R1"))
stab <- test_stability(R_score, dt$Y,
                       S_e = as.integer(dt$S_e),
                       S_o = as.integer(dt$S_o_02km))
print(stab)
```

Published result:

| Y | KS stat | p-value | n_exp | n_obs |
|---|--------:|--------:|------:|------:|
| 0 |  0.0800 |  0.2039 |   805 |   229 |
| 1 |  0.0904 |  0.8943 |   204 |    43 |

Both p-values are large. Failure to reject supports stability: the R–Y
relationship inside the imagery boundary can be transported to the 0–2 km buffer.

## 3. Simulation setup

Both simulations share the same calibrated DGP and model specification.

```{r dgp-setup, eval = run_chunks}
y_levels <- c(0L, 1L)
alpha_e  <- prop.table(table(dt$Y[dt$S_e      == 1]))
alpha_o  <- prop.table(table(dt$Y[dt$S_o_02km == 1]))
p0       <- as.numeric(alpha_e)

cat("alpha_e (Y|S_e=1):      Y=0:", round(p0[1], 4), " Y=1:", round(p0[2], 4), "\n")
cat("alpha_o (Y|S_o_02km=1): Y=0:", round(alpha_o[1], 4), " Y=1:", round(alpha_o[2], 4), "\n")

# R|Y lookup — R1 only (R0 = 1 - R1 adds no information for binary Y)
R_mat     <- matrix(dt$R1, ncol = 1, dimnames = list(NULL, "R1"))
R_given_Y <- lapply(y_levels, function(y) R_mat[dt$Y == y, , drop = FALSE])

TAU_GRID <- c(0, 0.05, 0.10, 0.15, 0.20)
B        <- 50L    # paper uses 500
SEED     <- 42L

models <- list(
  Y   = list(model = "logit"),
  D   = list(model = "logit"),
  S_e = list(model = "logit"),
  S_o = list(model = "logit")
)
```

## 4. No experimental outcomes simulation

Y is entirely unobserved in the experimental sample (`S_e = 1, S_o = 0`).

```{r noexp-run, eval = run_chunks}
run_noexp <- function(tau, b) {
  set.seed(SEED + b)
  p1  <- compute_p1_kl(p0, y_levels, tau)
  dat <- sim_noexp(1000L, 1000L, y_levels, p0, p1, as.numeric(alpha_o), R_given_Y)
  fit <- cv.rsv(Y=dat$Y, D=dat$D, S_e=dat$S_e, S_o=dat$S_o, R=dat$R,
                y_levels=y_levels, yK=y_levels[1], models=models,
                nfolds=2, seed=NULL, num.threads=1)
  fit <- add_se(fit, method = "influence")
  data.frame(tau=tau, tauhat=fit$coefficients, se=fit$se)
}

cat("Running no-experimental-outcomes simulation (B=50)...\n")
noexp_results <- do.call(rbind, lapply(TAU_GRID, function(tau) {
  cat(sprintf("  tau = %.2f\n", tau))
  do.call(rbind, lapply(seq_len(B), run_noexp, tau=tau))
}))
```

```{r noexp-summary, eval = run_chunks}
noexp_tbl <- do.call(rbind, lapply(TAU_GRID, function(tv)
  summarise_sims(noexp_results[noexp_results$tau == tv, ], tv)))
print(noexp_tbl)
```

## 5. No experimental outcomes: comparison to published results

```{r noexp-comparison, echo = FALSE, eval = TRUE}
pub_noexp <- data.frame(
  tau             = c(0.00, 0.05, 0.10, 0.15, 0.20),
  normalized_bias = c( 0.017,  0.024,  0.063,  0.000,  0.042),
  coverage        = c( 0.940,  0.950,  0.946,  0.950,  0.940),
  rmse            = c(0.0316, 0.0348, 0.0352, 0.0362, 0.0377),
  avg_se          = c(0.0318, 0.0330, 0.0341, 0.0354, 0.0363)
)
knitr::kable(
  pub_noexp,
  digits    = c(2, 3, 3, 4, 4),
  col.names = c("τ", "Norm. bias", "Coverage", "RMSE", "Avg SE"),
  caption   = paste("Published RSV results (no experimental outcomes):",
                    "n_e = 1,000, n_o = 1,000, 0–2 km, B = 500, B_se = 100.")
)
```

Near-zero normalized bias and ~94–95% coverage across all τ values confirm
that `cv.rsv()` is well-calibrated under this DGP.

## 6. With experimental outcomes simulation

A subset of `n_eo = 100` experimental units are also in the observational
sample (`S_e = S_o = 1`), so their outcomes are observed. This mirrors the
overlap design used throughout the package (see `?sim_rsv_data`).

```{r exp-run, eval = run_chunks}
N_EO <- 100L

run_exp <- function(tau, b) {
  set.seed(SEED + b)
  p1  <- compute_p1_kl(p0, y_levels, tau)
  dat <- sim_exp(1000L, 1000L, N_EO, y_levels, p0, p1, as.numeric(alpha_o), R_given_Y)
  fit <- cv.rsv(Y=dat$Y, D=dat$D, S_e=dat$S_e, S_o=dat$S_o, R=dat$R,
                y_levels=y_levels, yK=y_levels[1], models=models,
                nfolds=2, seed=NULL, num.threads=1)
  fit <- add_se(fit, method = "influence")
  data.frame(tau=tau, tauhat=fit$coefficients, se=fit$se)
}

cat("Running with-experimental-outcomes simulation (n_eo=100, B=50)...\n")
exp_results <- do.call(rbind, lapply(TAU_GRID, function(tau) {
  cat(sprintf("  tau = %.2f\n", tau))
  do.call(rbind, lapply(seq_len(B), run_exp, tau=tau))
}))
```

```{r exp-summary, eval = run_chunks}
exp_tbl <- do.call(rbind, lapply(TAU_GRID, function(tv)
  summarise_sims(exp_results[exp_results$tau == tv, ], tv)))
print(exp_tbl)
```

## 7. With experimental outcomes: comparison to published results

```{r exp-comparison, echo = FALSE, eval = TRUE}
pub_exp <- data.frame(
  tau             = c(0.00, 0.05, 0.10, 0.15, 0.20),
  normalized_bias = c( 0.017,  0.016,  0.060, -0.012,  0.020),
  coverage        = c( 0.932,  0.942,  0.928,  0.946,  0.938),
  rmse            = c(0.0317, 0.0345, 0.0350, 0.0361, 0.0370),
  avg_se          = c(0.0314, 0.0327, 0.0337, 0.0347, 0.0357)
)
knitr::kable(
  pub_exp,
  digits    = c(2, 3, 3, 4, 4),
  col.names = c("τ", "Norm. bias", "Coverage", "RMSE", "Avg SE"),
  caption   = paste("Published RSV results (with experimental outcomes):",
                    "n_e = 1,000, n_eo = 100, n_o = 1,000, 0–2 km, B = 500, B_se = 100.")
)
```

Results are similar to the no-experimental-outcomes case. Adding overlap units
(`n_eo = 100`) slightly reduces RMSE and average SE, consistent with the
efficiency gain from observing outcomes for some experimental units.

## 8. Reproducing the published results

To exactly match the published results, use `B = 500`, `score_bootstrap` with
`B_se = 100`, and parallel cores:

```{r full-replication, eval = FALSE}
n_cores <- min(100L, parallel::detectCores() - 1L)

run_full <- function(sim_fn, tau, b, ...) {
  set.seed(SEED + b)
  p1  <- compute_p1_kl(p0, y_levels, tau)
  dat <- sim_fn(p1 = p1, ...)
  fit <- cv.rsv(Y=dat$Y, D=dat$D, S_e=dat$S_e, S_o=dat$S_o, R=dat$R,
                y_levels=y_levels, yK=y_levels[1], models=models,
                nfolds=2, seed=NULL, num.threads=1)
  fit <- add_se(fit, method="score_bootstrap", B=100L, seed=NULL, num.threads=1)
  data.frame(tau=tau, tauhat=fit$coefficients, se=fit$se)
}

# No experimental outcomes (B=500)
noexp_full <- do.call(rbind, parallel::mclapply(
  seq_len(500L),
  function(b) do.call(rbind, lapply(TAU_GRID, run_full,
    sim_fn = function(p1, ...) sim_noexp(1000L, 1000L, y_levels, p0, p1,
                                         as.numeric(alpha_o), R_given_Y),
    b = b)),
  mc.cores = n_cores
))

# With experimental outcomes (B=500, n_eo=100)
exp_full <- do.call(rbind, parallel::mclapply(
  seq_len(500L),
  function(b) do.call(rbind, lapply(TAU_GRID, run_full,
    sim_fn = function(p1, ...) sim_exp(1000L, 1000L, 100L, y_levels, p0, p1,
                                       as.numeric(alpha_o), R_given_Y),
    b = b)),
  mc.cores = n_cores
))
```

Each full run takes approximately 30–60 minutes with 100 cores.
